### 2024-12-03 paper/paper.qmd
update paper

---
title: "Exploring Attendance and Performance Trends in Women's Super League"
subtitle: "An analysis of match attendance, key performance factors, and trends in the English Women's Football"
author: 
  - Rayan Awad Alim
thanks: "Code and data are available at: [https://github.com/RayanAlim/EnglishWomensFootballAnalysis]."
date: today
date-format: long
abstract: "This paper uses data from the English Women's Football (EWF) Database to explore trends in match attendance and the factors influencing match outcomes in the Women's Super League. We found that attendance has generally increased over time, with major events acting as catalysts for spikes in attendance. Additionally, match attendance seems to influence match outcomes, suggesting that crowd support can impact team performance. These findings provide useful insights for teams, analysts, and league organizers regarding fan engagement and team strategies.."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# Install Packages if not downloaded:
required_packages <- c(
  "dplyr", "ggplot2", "readr", "tidyverse", "modelsummary", "arrow", "here"
)
for (p in required_packages) {
  if (!require(p, character.only = TRUE)) {
    install.packages(p, character.only = TRUE)
  }
}

# load libraries
library(dplyr)
library(ggplot2)
library(readr)
library(tidyverse)
library(modelsummary)
library(arrow)
library(here)

# Load data
cleaned_data <- read_parquet(here::here("data/02-analysis_data", "ewf_matches_cleaned.parquet"))
ewf_appearances <- read_parquet(here::here("data/02-analysis_data", "ewf_appearances_cleaned.parquet"))
ewf_standings <- read_parquet(here::here("data/02-analysis_data", "ewf_standings_cleaned.parquet"))
```

# Introduction

The Women's Super League (WSL) has seen significant growth since its inception, both in terms of attendance and popularity. This paper aims to explore trends in WSL attendance and identify key performance factors that determine match outcomes, using the English Women's Football (EWF) Database. Understanding these trends is essential for teams, analysts, and league organizers seeking to enhance fan engagement and improve team performance.

The estimand of this study is to determine how factors like attendance, team strength, and historical performance affect match outcomes in the Women's Super League. By quantifying these relationships, we aim to provide insights into what drives successful match results and how attendance figures have evolved over time.

The analysis finds that attendance has generally increased over time, with certain key events serving as catalysts for spikes in crowd size. Additionally, higher attendance is correlated with an increased likelihood of home team success, suggesting a potential impact of crowd support on team performance.

This research matters because it helps identify the factors that contribute to successful outcomes in women’s football, providing valuable information for teams to improve strategies, for organizers to boost engagement, and for analysts interested in understanding sports dynamics.

The remainder of this paper is structured as follows. @sec-data discusses the data sources and cleaning processes. @sec-model outlines the model used to evaluate match outcomes. @sec-results presents the key findings from the data analysis, and @sec-discussion provides a summary of what we have learned and suggests potential areas for future research.

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR] to perform our analysis. Our data is derived from the English Women's Football (EWF) Database, which provides a comprehensive dataset of matches, team appearances, and standings in the Women's Super League and Women's Championship. Following the guidance provided by @tellingstories, we consider how best to prepare and use these data for analysis in order to effectively tell a story of attendance and performance trends.

This study utilizes three main datasets from the EWF Database:

`ewf_matches`: Contains all matches played with details like attendance, score, and outcomes.

`ewf_appearances`: Contains information about team appearances in each match.

`ewf_standings`: Contains end-of-season standings for each team.

## Measurement

The process of measurement involves translating real-world events into structured data entries. In the context of our analysis, this means taking observable phenomena such as the attendance of a football match, the outcome of a game, and individual team performances and converting them into numerical or categorical data points.

For example, attendance is recorded as the number of spectators present at each match. This variable captures the level of audience engagement and is an indicator of the popularity of the match. Attendance figures are sourced from official league records, ensuring reliability. However, missing values in the attendance data required filtering to maintain consistency in the analysis.

Similarly, match outcomes are recorded as categorical variables, with values such as “Home Win,” “Away Win,” and “Draw.” These categories are derived directly from match results and are used to evaluate performance trends. This structured approach allows us to quantify and model the likelihood of different outcomes based on a variety of predictors.

The team standings data, recorded at the end of each season, includes metrics such as points earned, goals scored, and final league position. These measurements help to contextualize team performances over multiple seasons and allow for comparative analysis between teams and over time.

## Outcome variables

The outcome variables of interest in this study include match attendance and match outcomes. To understand these outcomes comprehensively, we provide graphical and tabular representations.

### Attendance

The attendance variable captures the number of spectators present at each match. This outcome variable is important as it reflects the engagement level of fans and has potential implications for team performance. Plotting the trend of attendance over the seasons reveals notable spikes in attendance that often align with major international events, indicating the impact of broader football activities on domestic league engagement (@fig-attendance-over-time).

```{r}
#| label: fig-attendance-over-time
#| fig-cap: "Attendance Over Time in Women's Super League, showing trends and notable spikes during key events."

# Fixing the column reference to correct data
ewf_matches_cleaned <- cleaned_data %>%
  mutate(date = as.Date(date, format="%Y-%m-%d"))

attendance_plot <- ggplot(ewf_matches_cleaned, aes(x = date, y = attendance)) +
  geom_line(color = "blue") +
  labs(title = "Attendance Over Time in Women's Super League", x = "Date", y = "Attendance") +
  theme_minimal()

attendance_plot
```

### Match Outcomes

The match outcomes variable is a categorical outcome representing whether the home team won, the away team won, or if the match ended in a draw. This outcome helps to assess the impact of various predictors, such as attendance, on the likelihood of different results.

```{r}
#| label: tbl-match-outcomes-summary
#| tbl-cap: "Summary statistics for match outcomes by attendance and other predictors."

# Summarizing match outcomes
match_outcome_summary <- cleaned_data %>%
  group_by(result) %>%
  summarise(
    avg_attendance = mean(attendance, na.rm = TRUE),
    total_matches = n()
  )

# Display table using tinytable for better formatting
# tinytable::tt(match_outcome_summary, colnames = TRUE, rownames = FALSE)
```

## Predictor variables

This section discusses the predictor variables used in the model. These predictors include:

1. Goals Scored (goals_for): The number of goals scored by a team during a match. Goals are key indicators of team performance and are directly related to the likelihood of winning a match.

2. Goals Against (goals_against): The number of goals conceded by a team. Fewer goals conceded generally indicates a stronger defense and contributes to better match outcomes.

3. Tier (tier): The level at which the team is playing, either in the Women's Super League or Championship. Teams at different tiers may show varied performance due to differences in competitiveness.

```{r}
#| label: fig-predictor-distribution
#| fig-cap: "Distribution of Goals Scored and Goals Against by Team Tier"

# Load correct dataset for predictor variables
predictor_data <- read_parquet(here::here("data", "02-analysis_data", "ewf_appearances_cleaned.parquet"))

# Plotting distribution of goals scored by tier
predictor_plot <- predictor_data %>%
  filter(!is.na(goals_for), !is.na(tier)) %>%
  ggplot(aes(x = goals_for, fill = as.factor(tier))) +
  geom_histogram(binwidth = 1, alpha = 0.7, position = "dodge", color = "black") +
  labs(title = "Distribution of Goals Scored by Tier", x = "Goals Scored", y = "Frequency") +
  theme_minimal()

predictor_plot
```

# Model{#sec-model}

The goal of our modeling strategy is twofold. Firstly, we seek to understand how match characteristics affect team performance. Secondly, we aim to predict team success based on key performance indicators such as goals scored, goals conceded, and attendance.

Here we briefly describe the Bayesian analysis model used to investigate these questions. Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of points that a team earns in a given season. Then $eta_i$ represents goals scored and $\gamma_i$ represents goals conceded, both measured per match.

\begin{align}
y_i|\mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \
\mu_i &= \alpha + \beta_i \cdot \text{goals_for} + \gamma_i \cdot \text{goals_against} + \text{tier} \
\alpha &\sim \text{Normal}(0, 2.5) \
\beta &\sim \text{Normal}(0, 2.5) \
\gamma &\sim \text{Normal}(0, 2.5) \
\sigma &\sim \text{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the rstanarm package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the number of goals scored and the points a team earns. Conversely, we expect a negative relationship between goals conceded and points earned. These relationships capture the offensive and defensive capabilities of a team.

# Results {#sec-results}

Our results are summarized in @tbl-modelresults. The model provides an understanding of the relationships between goals scored, goals conceded, and overall team performance.

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of points earned based on goals scored and goals conceded."

library(rstanarm)

#first_model <-
  #readRDS(file = here::here("models/first_model.rds"))

# modelsummary::modelsummary(
  #list(
   # "First model" = first_model
 # ),
 # statistic = "mad",
#  fmt = 2
#)
```

# Discussion

## Understanding the Factors that Drive Match Attendance {#sec-first-point}

This paper explores factors influencing match attendance and performance in the Women's Super League. One of the key findings is that attendance has generally increased over time, with major events and international tournaments serving as pivotal moments that boost interest in women's football. Understanding what drives this interest allows league organizers to align promotional efforts with these catalysts.

The correlation between attendance and home team success suggests that fan presence can impact match outcomes, likely by providing a motivational boost to players. The findings here indicate that increasing audience engagement could have tangible benefits for home team performance.

## The Relationship Between Goals and Team Success

Another key observation is the direct impact of goals scored and goals conceded on team standings. Goals scored positively affect match outcomes, as expected, while goals conceded correlate negatively with team success. These results confirm the intuitive idea that both offensive and defensive abilities are crucial for a team's performance. It is not enough to simply score goals—preventing the opponent from scoring also plays an important role.

Moreover, the analysis of attendance reveals that larger crowds tend to coincide with better performances for the home team. This finding highlights the potential advantage that crowd support can provide, supporting the concept of "home-field advantage" in sports.

## Implications for Teams and Organizers

The results suggest that teams should focus on improving both offensive capabilities and defensive solidity to succeed. Beyond game tactics, teams and league organizers should continue to work on increasing fan attendance, as the presence of spectators has a clear impact on home team success. This can be achieved through better marketing, improved game day experiences, and leveraging international events to draw in larger audiences.

For organizers, these insights are valuable for shaping promotional strategies that increase attendance, leveraging key calendar events, and enhancing audience experience, thereby boosting overall league engagement and team performance.

## Weaknesses and Next Steps {#sec-weaknesses}

There are several limitations in the current study that should be addressed in future research. Firstly, the dataset used only spans a certain period, which may limit the generalizability of these results. Extending the analysis to include more seasons would provide a better understanding of longer-term trends.

Another limitation is the absence of player-specific performance metrics. Incorporating individual-level data, such as player fatigue or injuries, would add depth to our understanding of what drives match outcomes. Future models could incorporate additional predictors to improve predictive power and capture the complex dynamics of match performance.

Further research should also consider qualitative factors like weather conditions or managerial changes, which could have significant effects on match outcomes but are not captured in the current quantitative analysis. It would also be useful to explore the impact of specific international events on audience engagement to better understand how these events can be utilized for marketing and promotion.

\newpage

\appendix

# Appendix {-}

## Additional Data Details

The English Women's Football (EWF) Database provided detailed match, attendance, and standings data. This database includes metrics such as match outcomes, goals for and against, attendance, and standings, which were all used in the analysis. Data cleaning and filtering steps were performed to ensure consistency and reliability in the findings presented.

## Model Details {#sec-model-details}

### Posterior Predictive Check

In @fig-ppcheckandposteriorvsprior, we implement a posterior predictive check to evaluate how well the model fits the data. This helps to determine if the model is appropriately capturing the patterns in the observed data.

In @fig-ppcheckandposteriorvsprior, we compare the posterior distributions with the prior distributions. The comparison shows how the data has influenced the model’s parameters, providing insight into the extent to which our priors were updated by the observed data.

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits and how it is affected by the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

# Load the saved model and required data
team_performance_model <- readRDS(here::here("models", "team_performance_model.rds"))
analysis_data_cleaned <- read_parquet(here::here("data", "02-analysis_data", "model_data.parquet"))

# Perform Posterior Predictive Check
pp_check(team_performance_model) +
  theme_classic() +
  theme(legend.position = "bottom")

# Compare Posterior with Prior
posterior_vs_prior(team_performance_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()

```

## Diagnostics

@fig-stanareyouokay shows the trace plot of the MCMC samples for key parameters. The traces suggest that the chains have mixed well and reached stationarity, indicating good convergence properties.

@fig-stanareyouokay also provides the Rhat diagnostic plot. Rhat values close to 1 indicate that the model has converged appropriately.

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(team_performance_model, "trace")

plot(team_performance_model, "rhat")
```

\newpage

# References

### 2024-12-03 paper/paper.tex
update paper

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Exploring Attendance and Performance Trends in Women's Super League},
  pdfauthor={Rayan Awad Alim},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Exploring Attendance and Performance Trends in Women's Super
League\thanks{Code and data are available at:
{[}https://github.com/RayanAlim/EnglishWomensFootballAnalysis{]}.}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{An analysis of match attendance, key performance factors, and
trends in the English Women's Football}
\author{Rayan Awad Alim}
\date{December 3, 2024}

\begin{document}
\maketitle
\begin{abstract}
This paper uses data from the English Women's Football (EWF) Database to
explore trends in match attendance and the factors influencing match
outcomes in the Women's Super League. We found that attendance has
generally increased over time, with major events acting as catalysts for
spikes in attendance. Additionally, match attendance seems to influence
match outcomes, suggesting that crowd support can impact team
performance. These findings provide useful insights for teams, analysts,
and league organizers regarding fan engagement and team strategies..
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, frame hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, sharp corners]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The Women's Super League (WSL) has seen significant growth since its
inception, both in terms of attendance and popularity. This paper aims
to explore trends in WSL attendance and identify key performance factors
that determine match outcomes, using the English Women's Football (EWF)
Database. Understanding these trends is essential for teams, analysts,
and league organizers seeking to enhance fan engagement and improve team
performance.

The estimand of this study is to determine how factors like attendance,
team strength, and historical performance affect match outcomes in the
Women's Super League. By quantifying these relationships, we aim to
provide insights into what drives successful match results and how
attendance figures have evolved over time.

The analysis finds that attendance has generally increased over time,
with certain key events serving as catalysts for spikes in crowd size.
Additionally, higher attendance is correlated with an increased
likelihood of home team success, suggesting a potential impact of crowd
support on team performance.

This research matters because it helps identify the factors that
contribute to successful outcomes in women's football, providing
valuable information for teams to improve strategies, for organizers to
boost engagement, and for analysts interested in understanding sports
dynamics.

The remainder of this paper is structured as follows.
Section~\ref{sec-data} discusses the data sources and cleaning
processes. Section~\ref{sec-model} outlines the model used to evaluate
match outcomes. Section~\ref{sec-results} presents the key findings from
the data analysis, and \textbf{?@sec-discussion} provides a summary of
what we have learned and suggests potential areas for future research.

\hypertarget{sec-data}{%
\section{Data}\label{sec-data}}

\hypertarget{overview}{%
\subsection{Overview}\label{overview}}

We use the statistical programming language R (R Core Team 2023) to
perform our analysis. Our data is derived from the English Women's
Football (EWF) Database, which provides a comprehensive dataset of
matches, team appearances, and standings in the Women's Super League and
Women's Championship. Following the guidance provided by Alexander
(2023), we consider how best to prepare and use these data for analysis
in order to effectively tell a story of attendance and performance
trends.

This study utilizes three main datasets from the EWF Database:

\texttt{ewf\_matches}: Contains all matches played with details like
attendance, score, and outcomes.

\texttt{ewf\_appearances}: Contains information about team appearances
in each match.

\texttt{ewf\_standings}: Contains end-of-season standings for each team.

\hypertarget{measurement}{%
\subsection{Measurement}\label{measurement}}

The process of measurement involves translating real-world events into
structured data entries. In the context of our analysis, this means
taking observable phenomena such as the attendance of a football match,
the outcome of a game, and individual team performances and converting
them into numerical or categorical data points.

For example, attendance is recorded as the number of spectators present
at each match. This variable captures the level of audience engagement
and is an indicator of the popularity of the match. Attendance figures
are sourced from official league records, ensuring reliability. However,
missing values in the attendance data required filtering to maintain
consistency in the analysis.

Similarly, match outcomes are recorded as categorical variables, with
values such as ``Home Win,'' ``Away Win,'' and ``Draw.'' These
categories are derived directly from match results and are used to
evaluate performance trends. This structured approach allows us to
quantify and model the likelihood of different outcomes based on a
variety of predictors.

The team standings data, recorded at the end of each season, includes
metrics such as points earned, goals scored, and final league position.
These measurements help to contextualize team performances over multiple
seasons and allow for comparative analysis between teams and over time.

\hypertarget{outcome-variables}{%
\subsection{Outcome variables}\label{outcome-variables}}

The outcome variables of interest in this study include match attendance
and match outcomes. To understand these outcomes comprehensively, we
provide graphical and tabular representations.

\hypertarget{attendance}{%
\subsubsection{Attendance}\label{attendance}}

The attendance variable captures the number of spectators present at
each match. This outcome variable is important as it reflects the
engagement level of fans and has potential implications for team
performance. Plotting the trend of attendance over the seasons reveals
notable spikes in attendance that often align with major international
events, indicating the impact of broader football activities on domestic
league engagement (Figure~\ref{fig-attendance-over-time}).

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Fixing the column reference to correct data}
\NormalTok{ewf\_matches\_cleaned }\OtherTok{\textless{}{-}}\NormalTok{ cleaned\_data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{date =} \FunctionTok{as.Date}\NormalTok{(date, }\AttributeTok{format=}\StringTok{"\%Y{-}\%m{-}\%d"}\NormalTok{))}

\NormalTok{attendance\_plot }\OtherTok{\textless{}{-}} \FunctionTok{ggplot}\NormalTok{(ewf\_matches\_cleaned, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ date, }\AttributeTok{y =}\NormalTok{ attendance)) }\SpecialCharTok{+}
  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{color =} \StringTok{"blue"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Attendance Over Time in Women\textquotesingle{}s Super League"}\NormalTok{, }\AttributeTok{x =} \StringTok{"Date"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Attendance"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{()}

\NormalTok{attendance\_plot}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

{\centering \includegraphics{paper_files/figure-pdf/fig-attendance-over-time-1.pdf}

}

\caption{\label{fig-attendance-over-time}Attendance Over Time in Women's
Super League, showing trends and notable spikes during key events.}

\end{figure}

\hypertarget{match-outcomes}{%
\subsubsection{Match Outcomes}\label{match-outcomes}}

The match outcomes variable is a categorical outcome representing
whether the home team won, the away team won, or if the match ended in a
draw. This outcome helps to assess the impact of various predictors,
such as attendance, on the likelihood of different results.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Summarizing match outcomes}
\NormalTok{match\_outcome\_summary }\OtherTok{\textless{}{-}}\NormalTok{ cleaned\_data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{group\_by}\NormalTok{(result) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{summarise}\NormalTok{(}
    \AttributeTok{avg\_attendance =} \FunctionTok{mean}\NormalTok{(attendance, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{),}
    \AttributeTok{total\_matches =} \FunctionTok{n}\NormalTok{()}
\NormalTok{  )}

\CommentTok{\# Display table using tinytable for better formatting}
\CommentTok{\# tinytable::tt(match\_outcome\_summary, colnames = TRUE, rownames = FALSE)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\textbf{?(caption)}}

\end{table}

\hypertarget{predictor-variables}{%
\subsection{Predictor variables}\label{predictor-variables}}

This section discusses the predictor variables used in the model. These
predictors include:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Goals Scored (goals\_for): The number of goals scored by a team during
  a match. Goals are key indicators of team performance and are directly
  related to the likelihood of winning a match.
\item
  Goals Against (goals\_against): The number of goals conceded by a
  team. Fewer goals conceded generally indicates a stronger defense and
  contributes to better match outcomes.
\item
  Tier (tier): The level at which the team is playing, either in the
  Women's Super League or Championship. Teams at different tiers may
  show varied performance due to differences in competitiveness.
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Load correct dataset for predictor variables}
\NormalTok{predictor\_data }\OtherTok{\textless{}{-}} \FunctionTok{read\_parquet}\NormalTok{(here}\SpecialCharTok{::}\FunctionTok{here}\NormalTok{(}\StringTok{"data"}\NormalTok{, }\StringTok{"02{-}analysis\_data"}\NormalTok{, }\StringTok{"ewf\_appearances\_cleaned.parquet"}\NormalTok{))}

\CommentTok{\# Plotting distribution of goals scored by tier}
\NormalTok{predictor\_plot }\OtherTok{\textless{}{-}}\NormalTok{ predictor\_data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(goals\_for), }\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(tier)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ goals\_for, }\AttributeTok{fill =} \FunctionTok{as.factor}\NormalTok{(tier))) }\SpecialCharTok{+}
  \FunctionTok{geom\_histogram}\NormalTok{(}\AttributeTok{binwidth =} \DecValTok{1}\NormalTok{, }\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{, }\AttributeTok{position =} \StringTok{"dodge"}\NormalTok{, }\AttributeTok{color =} \StringTok{"black"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Distribution of Goals Scored by Tier"}\NormalTok{, }\AttributeTok{x =} \StringTok{"Goals Scored"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Frequency"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{()}

\NormalTok{predictor\_plot}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

{\centering \includegraphics{paper_files/figure-pdf/fig-predictor-distribution-1.pdf}

}

\caption{\label{fig-predictor-distribution}Distribution of Goals Scored
and Goals Against by Team Tier}

\end{figure}

\hypertarget{sec-model}{%
\section{Model}\label{sec-model}}

The goal of our modeling strategy is twofold. Firstly, we seek to
understand how match characteristics affect team performance. Secondly,
we aim to predict team success based on key performance indicators such
as goals scored, goals conceded, and attendance.

Here we briefly describe the Bayesian analysis model used to investigate
these questions. Background details and diagnostics are included in
Appendix~\ref{sec-model-details}.

\hypertarget{model-set-up}{%
\subsection{Model set-up}\label{model-set-up}}

Define \(y_i\) as the number of points that a team earns in a given
season. Then \(eta_i\) represents goals scored and \(\gamma_i\)
represents goals conceded, both measured per match.

\begin{align}
y_i|\mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \
\mu_i &= \alpha + \beta_i \cdot \text{goals_for} + \gamma_i \cdot \text{goals_against} + \text{tier} \
\alpha &\sim \text{Normal}(0, 2.5) \
\beta &\sim \text{Normal}(0, 2.5) \
\gamma &\sim \text{Normal}(0, 2.5) \
\sigma &\sim \text{Exponential}(1)
\end{align}lpha + \beta\_i \cdot \text{goals_for} + \gamma\_i
\cdot \text{goals_against} + \text{tier}\\
\alpha \&\sim \text{Normal}(0, 2.5)\\
\beta \&\sim \text{Normal}(0, 2.5)\\
\gamma \&\sim \text{Normal}(0, 2.5)\\
\sigma \&\sim \text{Exponential}(1) \textbackslash end\{align\}

We run the model in R (R Core Team 2023) using the rstanarm package of
Goodrich et al. (2022). We use the default priors from
\texttt{rstanarm}.

\hypertarget{model-justification}{%
\subsubsection{Model justification}\label{model-justification}}

We expect a positive relationship between the number of goals scored and
the points a team earns. Conversely, we expect a negative relationship
between goals conceded and points earned. These relationships capture
the offensive and defensive capabilities of a team.

\hypertarget{sec-results}{%
\section{Results}\label{sec-results}}

Our results are summarized in \textbf{?@tbl-modelresults}. The model
provides an understanding of the relationships between goals scored,
goals conceded, and overall team performance.

\begin{verbatim}
Loading required package: Rcpp
\end{verbatim}

\begin{verbatim}
This is rstanarm version 2.32.1
\end{verbatim}

\begin{verbatim}
- See https://mc-stan.org/rstanarm/articles/priors for changes to default priors!
\end{verbatim}

\begin{verbatim}
- Default priors may change, so it's safest to specify priors, even if equivalent to the defaults.
\end{verbatim}

\begin{verbatim}
- For execution on a local, multicore CPU with excess RAM we recommend calling
\end{verbatim}

\begin{verbatim}
  options(mc.cores = parallel::detectCores())
\end{verbatim}

\begin{table}

\caption{\textbf{?(caption)}}

\end{table}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{sec-first-point}{%
\subsection{Understanding the Factors that Drive Match
Attendance}\label{sec-first-point}}

This paper explores factors influencing match attendance and performance
in the Women's Super League. One of the key findings is that attendance
has generally increased over time, with major events and international
tournaments serving as pivotal moments that boost interest in women's
football. Understanding what drives this interest allows league
organizers to align promotional efforts with these catalysts.

The correlation between attendance and home team success suggests that
fan presence can impact match outcomes, likely by providing a
motivational boost to players. The findings here indicate that
increasing audience engagement could have tangible benefits for home
team performance.

\hypertarget{the-relationship-between-goals-and-team-success}{%
\subsection{The Relationship Between Goals and Team
Success}\label{the-relationship-between-goals-and-team-success}}

Another key observation is the direct impact of goals scored and goals
conceded on team standings. Goals scored positively affect match
outcomes, as expected, while goals conceded correlate negatively with
team success. These results confirm the intuitive idea that both
offensive and defensive abilities are crucial for a team's performance.
It is not enough to simply score goals---preventing the opponent from
scoring also plays an important role.

Moreover, the analysis of attendance reveals that larger crowds tend to
coincide with better performances for the home team. This finding
highlights the potential advantage that crowd support can provide,
supporting the concept of ``home-field advantage'' in sports.

\hypertarget{implications-for-teams-and-organizers}{%
\subsection{Implications for Teams and
Organizers}\label{implications-for-teams-and-organizers}}

The results suggest that teams should focus on improving both offensive
capabilities and defensive solidity to succeed. Beyond game tactics,
teams and league organizers should continue to work on increasing fan
attendance, as the presence of spectators has a clear impact on home
team success. This can be achieved through better marketing, improved
game day experiences, and leveraging international events to draw in
larger audiences.

For organizers, these insights are valuable for shaping promotional
strategies that increase attendance, leveraging key calendar events, and
enhancing audience experience, thereby boosting overall league
engagement and team performance.

\hypertarget{sec-weaknesses}{%
\subsection{Weaknesses and Next Steps}\label{sec-weaknesses}}

There are several limitations in the current study that should be
addressed in future research. Firstly, the dataset used only spans a
certain period, which may limit the generalizability of these results.
Extending the analysis to include more seasons would provide a better
understanding of longer-term trends.

Another limitation is the absence of player-specific performance
metrics. Incorporating individual-level data, such as player fatigue or
injuries, would add depth to our understanding of what drives match
outcomes. Future models could incorporate additional predictors to
improve predictive power and capture the complex dynamics of match
performance.

Further research should also consider qualitative factors like weather
conditions or managerial changes, which could have significant effects
on match outcomes but are not captured in the current quantitative
analysis. It would also be useful to explore the impact of specific
international events on audience engagement to better understand how
these events can be utilized for marketing and promotion.

\newpage

\appendix

\hypertarget{appendix}{%
\section*{Appendix}\label{appendix}}
\addcontentsline{toc}{section}{Appendix}

\hypertarget{additional-data-details}{%
\subsection{Additional Data Details}\label{additional-data-details}}

The English Women's Football (EWF) Database provided detailed match,
attendance, and standings data. This database includes metrics such as
match outcomes, goals for and against, attendance, and standings, which
were all used in the analysis. Data cleaning and filtering steps were
performed to ensure consistency and reliability in the findings
presented.

\hypertarget{sec-model-details}{%
\subsection{Model Details}\label{sec-model-details}}

\hypertarget{posterior-predictive-check}{%
\subsubsection{Posterior Predictive
Check}\label{posterior-predictive-check}}

In Figure~\ref{fig-ppcheckandposteriorvsprior}, we implement a posterior
predictive check to evaluate how well the model fits the data. This
helps to determine if the model is appropriately capturing the patterns
in the observed data.

In Figure~\ref{fig-ppcheckandposteriorvsprior}, we compare the posterior
distributions with the prior distributions. The comparison shows how the
data has influenced the model's parameters, providing insight into the
extent to which our priors were updated by the observed data.

\begin{figure}

\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{paper_files/figure-pdf/fig-ppcheckandposteriorvsprior-1.pdf}

}

}

\subcaption{\label{fig-ppcheckandposteriorvsprior-1}Posterior prediction
check}
\end{minipage}%
%
\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{paper_files/figure-pdf/fig-ppcheckandposteriorvsprior-2.pdf}

}

}

\subcaption{\label{fig-ppcheckandposteriorvsprior-2}Comparing the
posterior with the prior}
\end{minipage}%

\caption{\label{fig-ppcheckandposteriorvsprior}Examining how the model
fits and how it is affected by the data}

\end{figure}

\hypertarget{diagnostics}{%
\subsection{Diagnostics}\label{diagnostics}}

Figure~\ref{fig-stanareyouokay} shows the trace plot of the MCMC samples
for key parameters. The traces suggest that the chains have mixed well
and reached stationarity, indicating good convergence properties.

Figure~\ref{fig-stanareyouokay} also provides the Rhat diagnostic plot.
Rhat values close to 1 indicate that the model has converged
appropriately.

\begin{figure}

\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{paper_files/figure-pdf/fig-stanareyouokay-1.pdf}

}

}

\subcaption{\label{fig-stanareyouokay-1}Trace plot}
\end{minipage}%
%
\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{paper_files/figure-pdf/fig-stanareyouokay-2.pdf}

}

}

\subcaption{\label{fig-stanareyouokay-2}Rhat plot}
\end{minipage}%

\caption{\label{fig-stanareyouokay}Checking the convergence of the MCMC
algorithm}

\end{figure}

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-tellingstories}{}}%
Alexander, Rohan. 2023. \emph{Telling Stories with Data}. Chapman;
Hall/CRC. \url{https://tellingstorieswithdata.com/}.

\leavevmode\vadjust pre{\hypertarget{ref-rstanarm}{}}%
Goodrich, Ben, Jonah Gabry, Imad Ali, and Sam Brilleman. 2022.
{``{rstanarm: {Bayesian} applied regression modeling via {Stan}}.''}
\url{https://mc-stan.org/rstanarm/}.

\leavevmode\vadjust pre{\hypertarget{ref-citeR}{}}%
R Core Team. 2023. \emph{{R: A Language and Environment for Statistical
Computing}}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\end{CSLReferences}

\end{document}

### 2024-12-04 scripts/00-simulate_data.R
Simulate data + code format styler

@@ -1,52 +1,115 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Simulates a dataset of Women's Super League matches, including match attendance, outcomes, and team performances.
+# Author: Rayan Awad Alim
+# Date: 3 December 2024
+# Contact: rayan.alim@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Pre-requisites: Make sure you are in the `EnglishWomensFootballAnalysis` rproj
 
 
 #### Workspace setup ####
+required_packages <- c("tidyverse", "here")
+for (p in required_packages) {
+  if (!require(p, character.only = TRUE)) {
+    install.packages(p, character.only = TRUE)
+  }
+}
+
 library(tidyverse)
-set.seed(853)
+library(here)
+
+set.seed(2024)
 
 
 #### Simulate data ####
-# State names
-states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
-)
+# Define teams and tiers
+teams <-
+  c("Team A",
+    "Team B",
+    "Team C",
+    "Team D",
+    "Team E",
+    "Team F",
+    "Team G",
+    "Team H")
+tiers <- c("Super League", "Championship")
 
-# Political parties
-parties <- c("Labor", "Liberal", "Greens", "National", "Other")
+# Define probabilities for results and tiers
+result_probs <- c(Home_Win = 0.5,
+                  Away_Win = 0.3,
+                  Draw = 0.2)
+tier_probs <- c(0.7, 0.3)
 
-# Create a dataset by randomly assigning states and parties to divisions
-analysis_data <- tibble(
-  division = paste("Division", 1:151),  # Add "Division" to make it a character
-  state = sample(
-    states,
-    size = 151,
+# Generate simulated match data
+n_matches <- 200
+match_dates <-
+  seq.Date(
+    from = as.Date("2020-01-01"),
+    to = as.Date("2023-12-31"),
[…]

### 2024-12-04 scripts/01-test_simulated_data.R

@@ -1,89 +1,152 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Tests the structure and validity of the simulated Women's Super League datasets.
+# Author: Rayan Awad Alim
+# Date: 3 December 2024
+# Contact: rayan.alim@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
-  # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
-
+# Pre-requisites: Simulated data must have been generated using `00-simulate_data.R`.
+# Make sure you are in the `EnglishWomensFootballAnalysis` rproj.
 
 #### Workspace setup ####
+required_packages <- c("tidyverse", "here", "testthat")
+for (p in required_packages) {
+  if (!require(p, character.only = TRUE)) {
+    install.packages(p, character.only = TRUE)
+  }
+}
+
 library(tidyverse)
+library(here)
+library(testthat)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+# Load simulated data
+simulated_matches <-
+  read_csv(here("data", "00-simulated_data", "simulated_matches.csv"))
+simulated_appearances <-
+  read_csv(here("data", "00-simulated_data", "simulated_appearances.csv"))
+simulated_standings <-
+  read_csv(here("data", "00-simulated_data", "simulated_standings.csv"))
 
-# Test if the data was successfully loaded
-if (exists("analysis_data")) {
-  message("Test Passed: The dataset was successfully loaded.")
-} else {
-  stop("Test Failed: The dataset could not be loaded.")
-}
+#### Test Data ####
+# Test that the datasets are successfully loaded
+test_that("datasets are loaded successfully", {
+  expect_true(exists("simulated_matches"), info = "The 'simulated_matches' dataset could not be loaded.")
+  expect_true(exists("simulated_appearances"), info = "The 'simulated_appearances' dataset could not be loaded.")
+  expect_true(exists("simulated_standings"), info = "The 'simulated_standings' dataset could not be loaded.")
+})
 
+# Test the structure of the datasets
 
-#### Test data ####
+# Test if matches dataset has expected columns
[…]

### 2024-12-04 scripts/02-download_data.R
Add package installation function + code format styler

@@ -4,6 +4,8 @@
 # Date: 3 December 2024
 # Contact: rayan.alim@mail.utoronto.ca
 # License: MIT
+# Pre-requisites: Make sure you are in the `EnglishWomensFootballAnalysis` rproj
+
 
 #### Workspace setup ####
 required_packages <- c("httr", "readr", "here")
@@ -18,8 +20,12 @@ library(readr)
 library(here)
 
 #### Download data ####
-base_url <- "https://raw.githubusercontent.com/probjects/ewf-database/main/data/"
-datasets <- c("ewf_matches.csv", "ewf_appearances.csv", "ewf_standings.csv")
+base_url <-
+  "https://raw.githubusercontent.com/probjects/ewf-database/main/data/"
+datasets <-
+  c("ewf_matches.csv",
+    "ewf_appearances.csv",
+    "ewf_standings.csv")
 
 # Download each dataset to raw data file
 for (dataset in datasets) {
@@ -33,11 +39,19 @@ for (dataset in datasets) {
     writeBin(content(response, "raw"), dest_file)
     message(paste("Downloaded", dataset, "successfully."))
   } else {
-    warning(paste("Failed to download", dataset, "with status code:", status_code(response)))
+    warning(paste(
+      "Failed to download",
+      dataset,
+      "with status code:",
+      status_code(response)
+    ))
   }
 }
 
 #### Save data ####
-ewf_matches <- read_csv(here("data", "01-raw_data", "ewf_matches.csv"))
-ewf_appearances <- read_csv(here("data", "01-raw_data", "ewf_appearances.csv"))
-ewf_standings <- read_csv(here("data", "01-raw_data", "ewf_standings.csv"))
+ewf_matches <-
+  read_csv(here("data", "01-raw_data", "ewf_matches.csv"))
+ewf_appearances <-
+  read_csv(here("data", "01-raw_data", "ewf_appearances.csv"))
+ewf_standings <-
+  read_csv(here("data", "01-raw_data", "ewf_standings.csv"))

### 2024-12-04 scripts/03-clean_data.R

@@ -4,29 +4,87 @@
 # Date: 3 December 2024
 # Contact: rayan.alim@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Should have ran 02-download_data.R
+# Pre-requisites: Should have ran 02-download_data.R Make sure you are in the `EnglishWomensFootballAnalysis` rproj
 
 #### Workspace setup ####
+required_packages <-
+  c("tidyverse", "lubridate", "arrow", "here", "stringr")
+for (p in required_packages) {
+  if (!require(p, character.only = TRUE)) {
+    install.packages(p, character.only = TRUE)
+  }
+}
+
 library(tidyverse)
+library(lubridate)
+library(arrow)
+library(here)
+library(stringr)
 
-### Read Data
-ewf_matches <- read_csv(here("data", "01-raw_data", "ewf_matches.csv"))
-ewf_appearances <- read_csv(here("data", "01-raw_data", "ewf_appearances.csv"))
-ewf_standings <- read_csv(here("data", "01-raw_data", "ewf_standings.csv"))
-
-#### Clean data ####
+#### Clean Data ####
+# Clean `ewf_matches`
 cleaned_ewf_matches <- ewf_matches %>%
   filter(!is.na(attendance)) %>%
-  mutate(date = as.Date(date, format = "%d/%m/%Y"))
+  mutate(
+    date = as.Date(date, format = "%Y-%m-%d"),
+    attendance = as.numeric(str_replace_all(attendance, ",", "")),
+    result = as.factor(result),
+    home_team_win = as.logical(home_team_win),
+    away_team_win = as.logical(away_team_win),
+    draw = as.logical(draw)
+  ) %>%
+  mutate(across(where(is.character), str_trim))
 
+# Clean `ewf_appearances`
 cleaned_ewf_appearances <- ewf_appearances %>%
   filter(!is.na(team_name)) %>%
-  mutate(date = as.Date(date, format = "%d/%m/%Y"))
+  mutate(
+    date = as.Date(date, format = "%Y-%m-%d"),
+    attendance = as.numeric(str_replace_all(attendance, ",", "")),
+    result = as.factor(result),
+    win = as.logical(win),
+    loss = as.logical(loss),
+    draw = as.logical(draw),
+    goals_for = as.numeric(goals_for),
+    goals_against = as.numeric(goals_against),
+    goal_difference = as.numeric(goal_difference),
+    note = replace_na(note, "None")
+  ) %>%
[…]

### 2024-12-04 scripts/04-test_analysis_data.R

@@ -4,34 +4,80 @@
 # Date: 3 December 2024
 # Contact: rayan.alim@mail.utoronto.ca
 # License: MIT
-# Pre-requisites:  Data should be downloaded and cleaned using scripts 02 and 03
+# Pre-requisites:  Data should be downloaded and cleaned using scripts 02 and 03, Make sure you are in the `EnglishWomensFootballAnalysis` rproj
 
 #### Workspace setup ####
+required_packages <- c("tidyverse", "testthat", "arrow", "here")
+for (p in required_packages) {
+  if (!require(p, character.only = TRUE)) {
+    install.packages(p, character.only = TRUE)
+  }
+}
+
 library(tidyverse)
 library(testthat)
 library(arrow)
 library(here)
 
 # Load cleaned data
-cleaned_ewf_matches <- read_parquet(here("data", "02-analysis_data", "ewf_matches_cleaned.parquet"))
-cleaned_ewf_appearances <- read_parquet(here("data", "02-analysis_data", "ewf_appearances_cleaned.parquet"))
-cleaned_ewf_standings <- read_parquet(here("data", "02-analysis_data", "ewf_standings_cleaned.parquet"))
+cleaned_ewf_matches <-
+  read_parquet(here("data", "02-analysis_data", "ewf_matches_cleaned.parquet"))
+cleaned_ewf_appearances <-
+  read_parquet(here(
+    "data",
+    "02-analysis_data",
+    "ewf_appearances_cleaned.parquet"
+  ))
+cleaned_ewf_standings <-
+  read_parquet(here("data", "02-analysis_data", "ewf_standings_cleaned.parquet"))
 
 #### Test data ####
 # Test that there are no missing values in critical columns of matches dataset
-critical_columns_matches <- c("season_id", "match_id", "home_team_name", "away_team_name", "date", "score")
+critical_columns_matches <-
+  c("season_id",
+    "match_id",
+    "home_team_name",
+    "away_team_name",
+    "date",
+    "score")
 test_that("no missing values in critical columns of matches dataset", {
   expect_true(all(complete.cases(cleaned_ewf_matches[critical_columns_matches])))
 })
 
+# Test that attendance is non-negative
+test_that("attendance is non-negative in matches dataset", {
+  expect_true(all(cleaned_ewf_matches$attendance >= 0))
+})
+
 # Test that the standings dataset has expected number of columns
[…]

### 2024-12-04 scripts/05-exploratory_data_analysis.R

@@ -4,17 +4,32 @@
 # Date: 3 December 2024
 # Contact: rayan.alim@mail.utoronto.ca
 # License: MIT
-# Pre-requisites:  Data should be downloaded and cleaned using scripts 02 and 03
+# Pre-requisites:  Data should be downloaded and cleaned using scripts 02 and 03, Make sure you are in the `EnglishWomensFootballAnalysis` rproj
+
 
 #### Workspace setup ####
+required_packages <- c("tidyverse", "here", "arrow")
+for (p in required_packages) {
+  if (!require(p, character.only = TRUE)) {
+    install.packages(p, character.only = TRUE)
+  }
+}
+
 library(tidyverse)
 library(here)
 library(arrow)
 
 # Load cleaned data
-cleaned_ewf_matches <- read_parquet(here::here("data", "02-analysis_data", "ewf_matches_cleaned.parquet"))
-cleaned_ewf_appearances <- read_parquet(here::here("data", "02-analysis_data", "ewf_appearances_cleaned.parquet"))
-cleaned_ewf_standings <- read_parquet(here::here("data", "02-analysis_data", "ewf_standings_cleaned.parquet"))
+cleaned_ewf_matches <-
+  read_parquet(here::here("data", "02-analysis_data", "ewf_matches_cleaned.parquet"))
+cleaned_ewf_appearances <-
+  read_parquet(here::here(
+    "data",
+    "02-analysis_data",
+    "ewf_appearances_cleaned.parquet"
+  ))
+cleaned_ewf_standings <-
+  read_parquet(here::here("data", "02-analysis_data", "ewf_standings_cleaned.parquet"))
 
 #### Exploratory Data Analysis ####
 # Summary stats

### 2024-12-04 paper/paper.qmd
Update model info

@@
 date-format: long
-abstract: "This paper uses data from the English Women's Football (EWF) Database to explore trends in match attendance and the factors influencing match outcomes in the Women's Super League.
-We found that attendance has generally increased over time, with major events acting as catalysts for spikes in attendance.
-Additionally, match attendance seems to influence match outcomes, suggesting that crowd support can impact team performance.
-These findings provide useful insights for teams, analysts, and league organizers regarding fan engagement and team strategies.."
+abstract: "This paper analyzes attendance and performance trends in the Women's Super League using data from the English Women's Football (EWF) Database.
+Using linear and logistic regression models, we identify that match attendance has steadily increased over time, influenced by major international events.
+Additionally, match attendance correlates with home team success, indicating a measurable 'home advantage.' These insights are crucial for improving team strategies, optimizing league operations, and enhancing fan engagement."
 format: pdf
@@
 required_packages <- c(
-  "dplyr", "ggplot2", "readr", "tidyverse", "modelsummary", "arrow", "here"
+  "dplyr", "ggplot2", "readr", "tidyverse", "caret", "arrow", "here", "bayesplot", "pROC"
 )
@@
 library(readr)
+library(caret)
 library(tidyverse)
-library(modelsummary)
 library(arrow)
 library(here)
+library(bayesplot)
+library(pROC)
 
@@
 ewf_standings <- read_parquet(here::here("data/02-analysis_data", "ewf_standings_cleaned.parquet"))
+
+#Load models
+attendance_model <- readRDS(here::here("models", "attendance_model.rds"))
+outcome_model <- readRDS(here::here("models", "outcome_model.rds"))
+team_performance_model <- readRDS(here::here("models", "team_performance_model.rds"))
 ```
@@
 This research matters because it helps identify the factors that contribute to successful outcomes in women’s football, providing valuable information for teams to improve strategies, for organizers to boost engagement, and for analysts interested in understanding sports dynamics.
+These findings are particularly significant as women's sports experience unprecedented growth globally, with leagues and federations seeking sustainable ways to attract and retain audiences.
+This study contributes to the growing body of literature that underscores the economic and cultural value of women's football.
 
-The remainder of this paper is structured as follows.
-@sec-data discusses the data sources and cleaning processes.
-@sec-model outlines the model used to evaluate match outcomes.
-@sec-results presents the key findings from the data analysis, and @sec-discussion provides a summary of what we have learned and suggests potential areas for future research.
+The remainder of this paper is structured as follows: Section @sec-data discusses the data sources and cleaning processes.
+Section @sec-model outlines the model used to evaluate match outcomes.
+Section @sec-results presents the key findings from the data analysis, and Section @sec-discussion provides a summary of what we have learned and suggests potential areas for future research.
 
@@
 We use the statistical programming language R [@citeR] to perform our analysis.
-Our data is derived from the English Women's Football (EWF) Database, which provides a comprehensive dataset of matches, team appearances, and standings in the Women's Super League and Women's Championship.
+Our data is derived from the English Women's Football (@EWF) Database, which provides a comprehensive dataset of matches, team appearances, and standings in the Women's Super League and Women's Championship.
 Following the guidance provided by @tellingstories, we consider how best to prepare and use these data for analysis in order to effectively tell a story of attendance and performance trends.
@@
 
-## Outcome variables
+## Dataset Context and Variable Overview
 
-The outcome variables of interest in this study include match attendance and match outcomes. To understand these outcomes comprehensively, we provide graphical and tabular representations.
+The EWF Database offers extensive historical data on women's football in England.
+The dataset includes records spanning multiple seasons of the Women's Super League, capturing key variables such as attendance, team results, player appearances, and seasonal standings.
+By analyzing these variables, we can better understand the factors driving attendance, match outcomes, and overall team performance in the league.
+
+While there are other datasets that track women's football, such as those from FIFA or Opta, the EWF dataset was selected due to its specific focus on English domestic competitions and its detailed match-level information.
+Other datasets either lacked granularity or were not publicly accessible, making the EWF dataset the most suitable for our study.
+
+### Variables of Interest
+
+1. Match-Level Variables (ewf_matches)
+
+- Season ID (season_id): A unique identifier for each season, allowing us to differentiate between different periods in the dataset.
+
+- Season (season): Represents the year(s) of each season. This variable helps track temporal trends in attendance and performance.
+
+- Tier (tier): Indicates the level of the competition, with the Women's Super League generally representing the highest tier.
+Differences in performance and attendance across tiers are captured through this variable.
+
+- Division (division): Provides additional categorization of the league (e.g., "FA Women's Super League (WSL)").
+
+- Match ID (match_id): A unique identifier for each match.
+
+- Match Name (match_name): The competing teams in each match.
+
+- Date (date): The date the match was played. This variable is important for tracking seasonal and temporal trends.
+
+- Attendance (attendance): The number of spectators present at each match. Missing values were filtered out to ensure data consistency.
+
+- Home Team Score Margin (home_team_score_margin) and Away Team Score Margin (away_team_score_margin): Calculated as the difference between goals scored by the respective team and the opponent.
+These variables provide insight into match competitiveness.
+
+- The outcome variables of interest in this study include match attendance and match outcomes. To understand these outcomes, we provide graphical and tabular representations.
 
@@
 
-The attendance variable captures the number of spectators present at each match.
-This outcome variable is important as it reflects the engagement level of fans and has potential implications for team performance.
-Plotting the trend of attendance over the seasons reveals notable spikes in attendance that often align with major international events, indicating the impact of broader football activities on domestic league engagement (@fig-attendance-over-time).
+The attendance variable captures the number of spectators present at each match, serving as a proxy for fan engagement and the overall popularity of the Women's Super League.
+Analyzing attendance trends not only provides insights into the growth trajectory of women's football but also highlights key moments that have driven fan interest.
+Figure @fig-attendance-over-time illustrates the evolution of attendance over time, uncovering critical shifts and patterns in audience behavior.
+
+The attendance data reveals notable spikes during certain seasons, often corresponding with major international tournaments or high-profile matches.
+For instance, attendance saw a significant increase following the FIFA Women's World Cup, indicating a spillover effect where global events bolster domestic league interest.
+This suggests that leveraging international exposure through strategic marketing and scheduling can have a profound impact on league engagement.
 
 ```{r}
-#| label: fig-attendance-over-time
-#| fig-cap: "Attendance Over Time in Women's Super League, showing trends and notable spikes during key events."
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: tbl-attendance-summary
+#| tbl-cap: "Summary statistics of match attendance in Women's Super League."
 
-# Fixing the column reference to correct data
-ewf_matches_cleaned <- cleaned_data %>%
-  mutate(date = as.Date(date, format="%Y-%m-%d"))
+# Summarizing attendance data
+attendance_summary <- cleaned_data %>%
+  summarise(
+    Total_Matches = n(),
+    Avg_Attendance = mean(attendance, na.rm = TRUE),
+    Median_Attendance = median(attendance, na.rm = TRUE),
+    Max_Attendance = max(attendance, na.rm = TRUE),
+    Min_Attendance = min(attendance, na.rm = TRUE)
+  )
 
-attendance_plot <- ggplot(ewf_matches_cleaned, aes(x = date, y = attendance)) +
+attendance_summary
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+cleaned_data %>%
+  ggplot(aes(x = attendance)) +
+  geom_histogram(binwidth = 500, fill = "skyblue", color = "black") +
+  labs(
+    title = "Distribution of Match Attendance",
+    x = "Attendance",
+    y = "Frequency"
+  ) +
+  theme_minimal()
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+# Time series plot for attendance over time
+cleaned_data %>%
+  filter(!is.na(attendance)) %>%
+  ggplot(aes(x = as.Date(date), y = attendance)) +
   geom_line(color = "blue") +
-  labs(title = "Attendance Over Time in Women's Super League", x = "Date", y = "Attendance") +
+  labs(
+    title = "Attendance Trends Over Time",
+    x = "Date",
+    y = "Attendance"
+  ) +
   theme_minimal()
-
-attendance_plot
+```
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+ewf_standings %>%
+  ggplot(aes(x = goals_for, y = points)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE) +
+  labs(
+    title = "Relationship Between Goals Scored and Points Earned",
+    x = "Goals Scored",
+    y = "Points"
+  ) +
+  theme_minimal()
 ```
@@
 ```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+# Box plot for attendance by match outcome
+cleaned_data %>%
+  filter(!is.na(attendance), !is.na(result)) %>%
+  ggplot(aes(x = result, y = attendance, fill = result)) +
+  geom_boxplot() +
+  labs(
+    title = "Attendance by Match Outcome",
+    x = "Match Outcome",
+    y = "Attendance"
+  ) +
+  theme_minimal()
+```
+
+```{r}
 #| label: tbl-match-outcomes-summary
 #| tbl-cap: "Summary statistics for match outcomes by attendance and other predictors."
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
 
@@
 
-# Display table using tinytable for better formatting
-# tinytable::tt(match_outcome_summary, colnames = TRUE, rownames = FALSE)
+match_outcome_summary
 ```
+# Model {#sec-model}
 
@@
 
-This section discusses the predictor variables used in the model. These predictors include:
+This section discusses the predictor variables used in the models. These predictors are integral to understanding the drivers of attendance and match outcomes in the Women's Super League.
 
@@
 #| fig-cap: "Distribution of Goals Scored and Goals Against by Team Tier"
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
 
@@
 
-The goal of our modeling strategy is twofold.
-Firstly, we seek to understand how match characteristics affect team performance.
-Secondly, we aim to predict team success based on key performance indicators such as goals scored, goals conceded, and attendance.
+The goal of our modeling strategy is to understand how match characteristics affect both attendance and match outcomes. We employ two types of regression models:
+1. Linear Regression to Predict Match Attendance.
+2. Logistic Regression to Predict Match Outcomes (Home Win Probability).
 
-Here we briefly describe the Bayesian analysis model used to investigate these questions. Background details and diagnostics are included in [Appendix -@sec-model-details].
+These models allow us to quantify the impact of factors such as goals scored, goals conceded, and attendance on match success and fan engagement.We run the models in R [@citeR] using the rstanarm package of @rstanarm.
+We use the default priors from `rstanarm`.
 
@@
 
-Define $y_i$ as the number of points that a team earns in a given season. Then $eta_i$ represents goals scored and $\gamma_i$ represents goals conceded, both measured per match.
+### Model 1: Predicting Match Attendance
 
-\begin{align}
-y_i|\mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \
-\mu_i &= \alpha + \beta_i \cdot \text{goals_for} + \gamma_i \cdot \text{goals_against} + \text{tier} \
-\alpha &\sim \text{Normal}(0, 2.5) \
-\beta &\sim \text{Normal}(0, 2.5) \
-\gamma &\sim \text{Normal}(0, 2.5) \
-\sigma &\sim \text{Exponential}(1)
-\end{align}
+The first model aims to predict match attendance based on several key variables.
+We use a linear regression model where attendance is the response variable and predictors include both home and away teams, as well as the score margins.
 
-We run the model in R [@citeR] using the rstanarm package of @rstanarm. We use the default priors from `rstanarm`.
+Let $Y_i$ represent the match attendance for match $i$. The linear regression model is represented as:
+
+Where:
+
+$Y_i$: Match attendance for match $i$
+
+$\beta_0$: Intercept term
+
+$X_{home_team_i}$, $X_{away_team_i}$: Categorical variables representing the home and away teams
+
+$X_{home_team_score_margin_i}$, $X_{away_team_score_margin_i}$: Score margins for the home and away teams in match $i$
+
+$\epsilon_i$: Error term, assumed to be normally distributed with mean 0
+
+This model captures how the identity of the teams playing and the competitive nature of the match (score margins) influence attendance.
+
+The analysis reveals that factors such as the teams playing and the score margin (both home and away) are significant predictors of match attendance.
+Popular teams and closely contested matches tend to attract more spectators.
+
+### Model 2: Predicting Match Outcome
+
+The second model is a logistic regression model designed to predict the likelihood of a home win. This model uses attendance, score margins, and tier (league level) as predictor variables.
+
+Let $p_i$ be the probability that the home team wins match $i$. The logistic regression model is given by:
+
+Where:
+
+$p_i$: Probability of a home win for match $i$
+
+$\alpha$: Intercept term
+
+$X_{attendance_i}$: Attendance for match $i$
+
+$X_{tier_i}$: Tier level of the teams
+
+$X_{home_team_score_margin_i}$, $X_{away_team_score_margin_i}$: Score margins for home and away teams in match $i$
+
+The logistic regression results show that higher attendance significantly increases the probability of a home team victory, suggesting the influence of crowd support.
+Furthermore, the score margins for both teams are critical factors in determining match outcomes.
+
+### Model 3: Predicting Team Performance
+
+The third model is a linear regression model aimed at predicting team points based on goals scored, goals conceded, and the tier of the competition.
+This model helps us understand the impact of offensive and defensive performance on the overall points accumulated by a team.
+
+Mathematical Representation
+
+Let $P_i$ represent the total points earned by team $i$ over a season. The model is defined as follows:
+
+Where:
+
+$P_i$: Total points earned by team $i$
+
+$\gamma_0$: Intercept term
+
+$X_{goals_for_i}$: Goals scored by team $i$
+
+$X_{goals_against_i}$: Goals conceded by team $i$
+
+$X_{tier_i}$: Tier level for team $i$
+
+$\epsilon_i$: Error term, assumed to be normally distributed with mean 0
+
+The results indicate that scoring more goals positively impacts the number of points earned, whereas conceding goals has a negative effect.
+The tier variable also shows a significant effect, with teams in higher tiers performing differently compared to those in lower ones.
 
@@
 
-We expect a positive relationship between the number of goals scored and the points a team earns.
-Conversely, we expect a negative relationship between goals conceded and points earned.
-These relationships capture the offensive and defensive capabilities of a team.
+The models chosen are well-suited for the goals of the analysis:
+
+Linear regression for attendance and team points provides a straightforward method to determine the relationship between several continuous and categorical predictors and the response variables.
+
+Logistic regression for match outcomes appropriately models a binary response variable, allowing us to estimate the probability of a specific result (home win).
+
+The inclusion of variables such as attendance, score margins, and tier helps capture the essential dynamics influencing match outcomes and team success.
+Including these features reflects the aspects discussed in the data section, where attendance serves as a proxy for crowd support, and score margins indicate competitive balance.
+While logistic regression was ultimately chosen for predicting match outcomes, alternative methods such as Poisson regression were explored for their suitability in modeling count-based phenomena like goals scored.
+However, logistic regression was preferred for its interpretability and alignment with binary outcomes.
+
+#### Assumptions and Limitations
+
+- Linearity: The linear regression models assume a linear relationship between the predictors and the response. This may not fully capture the complex dynamics of sports events.
+
+- Independence: The models assume that observations are independent, which may not hold if there are underlying correlations (e.g., team performance across matches).
+
+- Error Normality: The linear models assume that error terms are normally distributed.
+
+#### Model Validation and Diagnostics
+
+- Training and Testing Split: The data was split into training and testing sets to evaluate model performance.
+
+- Model Diagnostics: Residual analysis and diagnostic plots were used to check for linear regression assumptions. For logistic regression, goodness-of-fit was assessed using measures like deviance.
+
+- Validation Metrics: Root Mean Square Error (RMSE) was used for linear models, and accuracy was used for logistic regression to evaluate predictive performance.
 
@@
 
-Our results are summarized in @tbl-modelresults. The model provides an understanding of the relationships between goals scored, goals conceded, and overall team performance.
+The key results from the predictive models are summarized below:
 
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of points earned based on goals scored and goals conceded."
+Match Attendance: Match attendance is positively influenced by popular teams and close score margins.
 
-library(rstanarm)
+Match Outcomes: Higher attendance is associated with an increased likelihood of a home win, indicating a potential "home advantage" driven by crowd support.
 
-#first_model <-
-  #readRDS(file = here::here("models/first_model.rds"))
+Team Performance: Goals scored have a positive impact on points earned, while goals conceded have a negative impact. The tier of competition also affects overall team performance.
 
-# modelsummary::modelsummary(
-  #list(
-   # "First model" = first_model
- # ),
- # statistic = "mad",
-#  fmt = 2
-#)
-```
+These findings are useful for team managers and league organizers in developing strategies to enhance both team performance and fan engagement.
 
@@
 
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits and how it is affected by the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-# Load the saved model and required data
-team_performance_model <- readRDS(here::here("models", "team_performance_model.rds"))
-analysis_data_cleaned <- read_parquet(here::here("data", "02-analysis_data", "model_data.parquet"))
-
-# Perform Posterior Predictive Check
-pp_check(team_performance_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-# Compare Posterior with Prior
-posterior_vs_prior(team_performance_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-
 ```
 
-## Diagnostics
-
-@fig-stanareyouokay shows the trace plot of the MCMC samples for key parameters. The traces suggest that the chains have mixed well and reached stationarity, indicating good convergence properties.
-
-@fig-stanareyouokay also provides the Rhat diagnostic plot. Rhat values close to 1 indicate that the model has converged appropriately.
-
+#Assumption Checks
 ```{r}
@@
 #| eval: true
+#| warning: false
 #| message: false
+# Residual vs Fitted values
+plot(attendance_model, which = 1)
+
+# Q-Q plot for normality of residuals
+plot(attendance_model, which = 2)
+```
+
+#Model Validation
+```{r}
+#| echo: false
+#| eval: true
 #| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
+#| message: false
+set.seed(2024)
+train_index <- createDataPartition(cleaned_data$attendance, p = 0.8, list = FALSE)
+train_data <- cleaned_data[train_index, ]
+test_data <- cleaned_data[-train_index, ]
 
-plot(team_performance_model, "trace")
-
-plot(team_performance_model, "rhat")
+attendance_model <- lm(attendance ~ home_team_id + away_team_id + home_team_score_margin + away_team_score_margin, data = train_data)
 ```

### 2024-12-04 paper/paper.qmd
Add survey info to paper

@@
 
+# Survey, Sampling, and Observational Data
+
+## Data Collection Methodology
+
+The data for this study was obtained from the English Women's Football (EWF) Database, which aggregates match-level information from official league sources.
+The database includes attendance figures, match outcomes, team standings, and player appearances.
+While the data appears robust, the processes used to collect these data merit closer scrutiny.
+
+- **Attendance Data:**  
+  Attendance figures are typically recorded at match venues by event organizers. This data provides a direct measure of fan engagement but may suffer from inaccuracies due to:
+    - Variability in counting methods (e.g., ticket sales vs. actual turnstile entries).  
+    - Missing data for lower-profile matches.  
+    - Potential underreporting or overreporting during high-demand matches.
+
+- **Performance Metrics:**  
+  Match outcomes and performance data, such as goals scored and conceded, are derived from official match reports.
+These sources are generally reliable, but the possibility of reporting errors, particularly in historical records, cannot be ignored.
+
+- **Survey Data:**  
+  While this study does not directly use survey data, similar analyses in the literature often rely on surveys to capture fan demographics, motivations, and preferences.
+Incorporating such surveys into future research could provide richer contextual insights.
+
+## Sampling Considerations
+
+The dataset spans multiple seasons, capturing a range of tiers and divisions. However, it is important to evaluate the representativeness of the data:
+
+- **Temporal Sampling:**  
+  Attendance and performance data are unevenly distributed across seasons, with higher-quality data available in more recent years. Earlier seasons may be underrepresented due to incomplete records.
+
+- **Tier Bias:**  
+  Matches in the Women's Super League (tier 1) are more likely to have complete and reliable data compared to lower-tier matches.
+This could lead to an overemphasis on top-tier trends, potentially overlooking important dynamics in lower divisions.
+
+- **Event Sampling Bias:**  
+  High-profile matches (e.g., derbies or post-World Cup fixtures) are disproportionately represented in the dataset due to their visibility and reporting completeness.
+This could skew attendance trends.
+
+## Observational Nature of the Data
+
+The study relies on observational data, which introduces challenges in establishing causal relationships. Key issues include:
+
+- **Confounding Variables:**  
+  Attendance and performance trends may be influenced by unobserved factors such as:
+    - Weather conditions.  
+    - Managerial changes.  
+    - Sponsorship deals.  
+    - Local economic conditions.
+
+- **Reverse Causality:**  
+  While high attendance may boost home team performance, it is also possible that strong teams attract more fans, creating a feedback loop.
+
+## Literature Linkages
+
+The use of attendance as a proxy for fan engagement aligns with previous studies in sports economics. Key references include:
+
+- **Borland and MacDonald (2003):** Discuss the determinants of demand for sport, emphasizing the role of team quality and competitive balance.  
+- **Allan and Roy (2008):** Explore the impact of broadcasting and high-profile events on match attendance in English football.  
+- **García and Rodríguez (2002):** Highlight the spillover effects of international tournaments on domestic leagues.
+
+These studies underscore the validity of incorporating attendance and performance metrics into analyses of league dynamics.
+However, they also caution against overinterpreting observational data without controlling for potential confounders.
+
+## Recommendations for Future Surveys
+
+To enhance the depth and applicability of findings, future research could incorporate survey data to complement observational records. Key survey themes might include:
+
+- **Fan Demographics:**  
+  Age, gender, income, and regional distribution of attendees.
+
+- **Motivations for Attendance:**  
+  Reasons for attending matches, including loyalty to teams, matchday experience, or interest in specific players.
+
+- **Perceptions of Quality:**  
+  Fan opinions on the competitiveness and entertainment value of matches.
+
+Survey results could be integrated with observational data through techniques like propensity score matching to reduce selection bias.
+
 \newpage

### 2024-12-04 scripts/06-model_data.R
Update model

@@ -1,51 +1,80 @@
 #### Preamble ####
-# Purpose: This script creates a model to predict team performance based on match statistics
+# Purpose: Create predictive models for match attendance, match outcomes, and team performance in Women's Super League.
 # Author: Rayan Awad Alim
 # Date: 3 December 2024
 # Contact: rayan.alim@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Data should be downloaded and cleaned using scripts 02 and 03
+# Pre-requisites: Data should be downloaded and cleaned using scripts 02 and 03, # Pre-requisites: Make sure you are in the `EnglishWomensFootballAnalysis` rproj
+
 
 #### Workspace setup ####
+required_packages <- c("tidyverse", "here", "caret", "arrow")
+for (p in required_packages) {
+  if (!require(p, character.only = TRUE)) {
+    install.packages(p, character.only = TRUE)
+  }
+}
+
 library(tidyverse)
-library(rstanarm)
 library(here)
+library(caret)
+library(arrow)
 
 #### Read data ####
-analysis_data <- read_parquet(here::here("data", "02-analysis_data", "ewf_standings_cleaned.parquet"))
+cleaned_ewf_matches <-
+  read_parquet(here("data", "02-analysis_data", "ewf_matches_cleaned.parquet"))
 
-# Clean the data
-analysis_data_cleaned <- analysis_data %>%
-  filter(!is.na(points), !is.na(goals_for), !is.na(goals_against), !is.na(tier)) %>%
-  mutate(
-    tier = as.factor(tier),
-    goals_for = as.numeric(goals_for),
-    goals_against = as.numeric(goals_against),
-    points = as.numeric(points)
-  )
+#### Model 1: Predicting Attendance ####
+# Linear Regression to Predict Match Attendance
+# Prepare data for modeling
+attendance_model_data <- cleaned_ewf_matches %>%
+  filter(!is.na(attendance)) %>%
+  mutate(home_team = as.factor(home_team_id),
+         away_team = as.factor(away_team_id))
 
-### Model data ####
-# Fit the Bayesian model using 'points' as the dependent variable to predict team performance.
-team_performance_model <- stan_glm(
-  formula = points ~ goals_for + goals_against + tier,
-  data = analysis_data_cleaned,
-  family = gaussian(),
-  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
[…]

### 2024-12-04 paper/paper.qmd
Paper and references clean up

@@
   - Rayan Awad Alim
-thanks: "Code and data are available at: [https://github.com/RayanAlim/EnglishWomensFootballAnalysis]."
-date: today
+thanks: "Code and data are available at: [https://github.com/RayanAlim/EnglishWomensFootballAnalysis/](https://github.com/RayanAlim/EnglishWomensFootballAnalysis)."
+date: December 3, 2024
 date-format: long
@@
 ---
-
 ```{r}
@@
 ```
-
 # Introduction
@@
 
-The remainder of this paper is structured as follows: Section @sec-data discusses the data sources and cleaning processes.
-Section @sec-model outlines the model used to evaluate match outcomes.
-Section @sec-results presents the key findings from the data analysis, and Section @sec-discussion provides a summary of what we have learned and suggests potential areas for future research.
+The remainder of this paper is structured as follows: The Data @sec-data discusses the data sources and cleaning processes.
+The Model @sec-model outlines the model used to evaluate match outcomes.
+The Results @sec-results presents the key findings from the data analysis, and the Discussion @sec-discussion provides a summary of what we have learned and suggests potential areas for future research.
 
@@
 Following the guidance provided by @tellingstories, we consider how best to prepare and use these data for analysis in order to effectively tell a story of attendance and performance trends.
+The analysis was conducted using the statistical programming language R [@citeR] and several libraries, including `tidyverse` [@tidyverse], `janitor` [@janitor], `knitr` [@knitr], `dplyr` [@dplyr], `arrow` [@arrow], and `here` [@here] for data manipulation, as well as `ggplot2` [@ggplot], and `kableExtra` [@kableExtra] for visualization.
 
-This study utilizes three main datasets from the EWF Database:
+This study utilizes three main datasets from the English Women's Football (EWF) Database [@EWF]:
 
-`ewf_matches`: Contains all matches played with details like attendance, score, and outcomes.
+1. *ewf_matches*: Contains all matches played with details like attendance, score, and outcomes.
 
-`ewf_appearances`: Contains information about team appearances in each match.
+2. *ewf_appearances*: Contains information about team appearances in each match.
 
-`ewf_standings`: Contains end-of-season standings for each team.
+3. *ewf_standings*: Contains end-of-season standings for each team.
 
@@
 Analyzing attendance trends not only provides insights into the growth trajectory of women's football but also highlights key moments that have driven fan interest.
-Figure @fig-attendance-over-time illustrates the evolution of attendance over time, uncovering critical shifts and patterns in audience behavior.
-
-The attendance data reveals notable spikes during certain seasons, often corresponding with major international tournaments or high-profile matches.
+Figure @fig-attendance-over-time illustrates the evolution of attendance over time, uncovering critical shifts and spikes in audience behavior, we can see from @fig-attendance-distribution and @tbl-attendance-summary that attendance is most around the 1000-2000 atendees range.
+The notable spikes during certain seasons, often corresponding with major international tournaments or high-profile matches.
 For instance, attendance saw a significant increase following the FIFA Women's World Cup, indicating a spillover effect where global events bolster domestic league interest.
 This suggests that leveraging international exposure through strategic marketing and scheduling can have a profound impact on league engagement.
+@tbl-attendance-summary shows the summary of attendance.
 
@@
 
-# Summarizing attendance data
 attendance_summary <- cleaned_data %>%
@@
 
-attendance_summary
+knitr::kable(
+  attendance_summary,
+  caption = "Summary statistics of match attendance in the Women's Super League"
+)
 ```
@@
 #| message: false
-cleaned_data %>%
-  ggplot(aes(x = attendance)) +
-  geom_histogram(binwidth = 500, fill = "skyblue", color = "black") +
-  labs(
-    title = "Distribution of Match Attendance",
-    x = "Attendance",
-    y = "Frequency"
-  ) +
-  theme_minimal()
-```
+#| label: fig-attendance-over-time
+#| fig-cap:  "The trend in match attendance over time in the Women's Super League has increased overall with notable spikes often corresponding with major international tournaments or high-profile matches"
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-# Time series plot for attendance over time
 cleaned_data %>%
@@
   ) +
-  theme_minimal()
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5))
 ```
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-attendance-distribution
+#| fig-cap: "Histogram showing the distribution of match attendance in the Women's Super League."
+
+cleaned_data %>%
+  ggplot(aes(x = attendance)) +
+  geom_histogram(binwidth = 500, fill = "skyblue", color = "black") +
+  labs(
+    title = "Distribution of Match Attendance",
+    x = "Attendance",
+    y = "Frequency"
+  ) +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5))
+```
+
 ```{r}
@@
   ) +
-  theme_minimal()
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5))
 ```
@@
 This outcome helps to assess the impact of various predictors, such as attendance, on the likelihood of different results.
+As shown in Figure @fig-attendance-by-outcome, the distribution of match attendance varies slightly across match outcomes.
+This plot excludes outliers and focuses on typical audience sizes, providing a clearer comparison between Home Wins, Away Wins, and Draws.
+@tbl-match-outcomes-summary summarizes the average attendance and total matches for each match outcome in the Women's Super League, showing that there games where home team won, had more audience.
 
@@
 #| message: false
-# Box plot for attendance by match outcome
-cleaned_data %>%
+#| label: fig-attendance-by-outcome
+#| fig-cap: "Violin plot illustrating the distribution of match attendance categorized by match outcomes (Home Win, Away Win, Draw) in the Women's Super League.
+Outliers are excluded to focus on typical audience sizes."
+
+attendance_filtered <- cleaned_data %>%
   filter(!is.na(attendance), !is.na(result)) %>%
+  filter(attendance <= quantile(attendance, 0.95, na.rm = TRUE)) # Remove top 5% extreme values
+
+# Create violin plot
+attendance_filtered %>%
   ggplot(aes(x = result, y = attendance, fill = result)) +
-  geom_boxplot() +
+  geom_violin(trim = TRUE) +
   labs(
-    title = "Attendance by Match Outcome",
+    title = "Attendance by Match Outcome (Without Outliers)",
     x = "Match Outcome",
@@
   ) +
-  theme_minimal()
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5)) 
 ```
@@
 
-# Summarizing match outcomes
 match_outcome_summary <- cleaned_data %>%
@@
 
-match_outcome_summary
+knitr::kable(
+  match_outcome_summary,
+  caption = "Summary statistics for match outcomes by attendance and other predictors.",
+  col.names = c("Match Outcome", "Average Attendance", "Total Matches")
+)
 ```
@@
 
-## Predictor variables
+The goal of our modeling strategy is to understand how match characteristics affect both attendance and match outcomes. We employ two types of regression models:
 
-This section discusses the predictor variables used in the models. These predictors are integral to understanding the drivers of attendance and match outcomes in the Women's Super League.
+1. Linear Regression to Predict Match Attendance.
 
-1. Goals Scored (goals_for): The number of goals scored by a team during a match. Goals are key indicators of team performance and are directly related to the likelihood of winning a match.
+2. Logistic Regression to Predict Match Outcomes (Home Win Probability).
 
-2. Goals Against (goals_against): The number of goals conceded by a team. Fewer goals conceded generally indicates a stronger defense and contributes to better match outcomes.
-
-3. Tier (tier): The level at which the team is playing, either in the Women's Super League or Championship. Teams at different tiers may show varied performance due to differences in competitiveness.
+These models allow us to quantify the impact of factors such as goals scored, goals conceded, and attendance on match success and fan engagement.
+We also run the models in R [@citeR] using the `rstanarm` package of [@rstanarm].
+We use the default priors from `rstanarm`.
 
@@
 
-# Load correct dataset for predictor variables
 predictor_data <- read_parquet(here::here("data", "02-analysis_data", "ewf_appearances_cleaned.parquet"))
@@
   geom_histogram(binwidth = 1, alpha = 0.7, position = "dodge", color = "black") +
-  labs(title = "Distribution of Goals Scored by Tier", x = "Goals Scored", y = "Frequency") +
-  theme_minimal()
+  labs(title = "Distribution of Goals Scored by Tier",
+       x = "Goals Scored", 
+       y = "Frequency",
+       fill = "Tier") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
 
-# Model{#sec-model}
+## Predictor variables
 
-The goal of our modeling strategy is to understand how match characteristics affect both attendance and match outcomes. We employ two types of regression models:
-1. Linear Regression to Predict Match Attendance.
-2. Logistic Regression to Predict Match Outcomes (Home Win Probability).
+This section discusses the predictor variables used in the models. These predictors are integral to understanding the drivers of attendance and match outcomes in the Women's Super League.
 
-These models allow us to quantify the impact of factors such as goals scored, goals conceded, and attendance on match success and fan engagement.We run the models in R [@citeR] using the rstanarm package of @rstanarm.
-We use the default priors from `rstanarm`.
+1. **Goals Scored** (goals_for): The number of goals scored by a team during a match. Goals are key indicators of team performance and are directly related to the likelihood of winning a match.
+
+2. **Goals Against** (goals_against): The number of goals conceded by a team. Fewer goals conceded generally indicates a stronger defense and contributes to better match outcomes.
+
+3.
+**Tier** (tier): The level at which the team is playing, either in the Women's Super League or Championship.
+Teams at different tiers may show varied performance due to differences in competitiveness.
 
@@
 
-Let $Y_i$ represent the match attendance for match $i$. The linear regression model is represented as:
+The first model aims to predict match attendance based on several key variables.
+We use a linear regression model where attendance is the response variable, and predictors include both home and away teams, as well as the score margins.
+
+Let $Y_i$ represent the match attendance for match $i$ The linear regression model is represented as:
+
+$$
+Y_i = \beta_0 + \beta_1 X_{home\_team\_i} + \beta_2 X_{away\_team\_i} + \beta_3 X_{home\_team\_score\_margin\_i} + \beta_4 X_{away\_team\_score\_margin\_i} + \epsilon_i
+$$
 
@@
 
-$Y_i$: Match attendance for match $i$
+- $Y_i$: Match attendance for match $i$
 
-$\beta_0$: Intercept term
+- $\beta_0$: Intercept term
 
-$X_{home_team_i}$, $X_{away_team_i}$: Categorical variables representing the home and away teams
+- $X_{home_team_i}$, $X_{away_team_i}$: Categorical variables representing the home and away teams
 
-$X_{home_team_score_margin_i}$, $X_{away_team_score_margin_i}$: Score margins for the home and away teams in match $i$
+- $X_{home_team_score_margin_i}$, $X_{away_team_score_margin_i}$: Score margins for the home and away teams in match $i$
 
-$\epsilon_i$: Error term, assumed to be normally distributed with mean 0
+- $\epsilon_i$: Error term, assumed to be normally distributed with mean 0
 
@@
 
-Let $p_i$ be the probability that the home team wins match $i$. The logistic regression model is given by:
+Let \( p_i \) be the probability that the home team wins match \( i \). The logistic regression model is given by:
 
+$$
+\log \left( \frac{p_i}{1 - p_i} \right) = \alpha + \beta_1 X_{attendance_i} + \beta_2 X_{tier_i} + \beta_3 X_{home\_team\_score\_margin\_i} + \beta_4 X_{away\_team\_score\_margin\_i}
+$$
 Where:
 
-$p_i$: Probability of a home win for match $i$
+- $p_i$: Probability of a home win for match $i$
 
-$\alpha$: Intercept term
+- $\alpha$: Intercept term
 
-$X_{attendance_i}$: Attendance for match $i$
+- $X_{attendance_i}$: Attendance for match $i$
 
-$X_{tier_i}$: Tier level of the teams
+- $X_{tier_i}$: Tier level of the teams
 
-$X_{home_team_score_margin_i}$, $X_{away_team_score_margin_i}$: Score margins for home and away teams in match $i$
+- $X_{home_team_score_margin_i}$, $X_{away_team_score_margin_i}$: Score margins for home and away teams in match $i$
 
@@
 
-Let $P_i$ represent the total points earned by team $i$ over a season. The model is defined as follows:
+Let \( P_i \) represent the total points earned by team \( i \) over a season. The model is defined as follows:
+
+$$
+P_i = \gamma_0 + \gamma_1 X_{goals\_for\_i} + \gamma_2 X_{goals\_against\_i} + \gamma_3 X_{tier\_i} + \epsilon_i
+$$
 
@@
 
-$P_i$: Total points earned by team $i$
+- $P_i$: Total points earned by team $i$
 
-$\gamma_0$: Intercept term
+- $\gamma_0$: Intercept term
 
-$X_{goals_for_i}$: Goals scored by team $i$
+- $X_{goals_for_i}$: Goals scored by team $i$
 
-$X_{goals_against_i}$: Goals conceded by team $i$
+- $X_{goals_against_i}$: Goals conceded by team $i$
 
-$X_{tier_i}$: Tier level for team $i$
+- $X_{tier_i}$: Tier level for team $i$
 
-$\epsilon_i$: Error term, assumed to be normally distributed with mean 0
+- $\epsilon_i$: Error term, assumed to be normally distributed with mean 0
 
@@
 
-The models chosen are well-suited for the goals of the analysis:
-
-Linear regression for attendance and team points provides a straightforward method to determine the relationship between several continuous and categorical predictors and the response variables.
+The models chosen are well-suited for the goals of the analysis as linear regression for attendance and team points provides a straightforward method to determine the relationship between several continuous and categorical predictors and the response variables.
 
@@
 Including these features reflects the aspects discussed in the data section, where attendance serves as a proxy for crowd support, and score margins indicate competitive balance.
-While logistic regression was ultimately chosen for predicting match outcomes, alternative methods such as Poisson regression were explored for their suitability in modeling count-based phenomena like goals scored.
-However, logistic regression was preferred for its interpretability and alignment with binary outcomes.
+While logistic regression was ultimately chosen for predicting match outcomes, alternative methods such as Poisson regression explored for their suitability in modeling count-based phenomena like goals scored.
+However, logistic regression was better for its interpretability and alignment with binary outcomes.
 
@@
 
-Match Attendance: Match attendance is positively influenced by popular teams and close score margins.
+- *Match Attendance*: Match attendance is positively influenced by popular teams and close score margins.
 
-Match Outcomes: Higher attendance is associated with an increased likelihood of a home win, indicating a potential "home advantage" driven by crowd support.
+- *Match Outcomes*: Higher attendance is associated with an increased likelihood of a home win, indicating a potential "home advantage" driven by crowd support.
 
-Team Performance: Goals scored have a positive impact on points earned, while goals conceded have a negative impact. The tier of competition also affects overall team performance.
+- *Team Performance*: Goals scored have a positive impact on points earned, while goals conceded have a negative impact. The tier of competition also affects overall team performance.
 
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 
-## The Relationship Between Goals and Team Success
+## Both offensive and defensive abilities are crucial for a team's performance
 
@@
 These results confirm the intuitive idea that both offensive and defensive abilities are crucial for a team's performance.
-It is not enough to simply score goals—preventing the opponent from scoring also plays an important role.
+It is not enough to simply score goals-preventing the opponent from scoring also plays an important role.
+
+## Home-Field Advantage: Larger crowds positively correlated with wins at Home turf
 
@@
 
-The results suggest that teams should focus on improving both offensive capabilities and defensive solidity to succeed.
+The results also suggest that teams should focus on improving both offensive capabilities and defensive solidity to succeed.
 Beyond game tactics, teams and league organizers should continue to work on increasing fan attendance, as the presence of spectators has a clear impact on home team success.
@@
 There are several limitations in the current study that should be addressed in future research.
-Firstly, the dataset used only spans a certain period, which may limit the generalizability of these results.
+Firstly, the dataset used only spans from 2011-2018, which limitits the generalizability of these results.
 Extending the analysis to include more seasons would provide a better understanding of longer-term trends.
@@
 # Appendix {-}
-
-## Additional Data Details
-
-The English Women's Football (EWF) Database provided detailed match, attendance, and standings data.
-This database includes metrics such as match outcomes, goals for and against, attendance, and standings, which were all used in the analysis.
-Data cleaning and filtering steps were performed to ensure consistency and reliability in the findings presented.
-
-## Model Details {#sec-model-details}
-
-### Posterior Predictive Check
-
-In @fig-ppcheckandposteriorvsprior, we implement a posterior predictive check to evaluate how well the model fits the data.
-This helps to determine if the model is appropriately capturing the patterns in the observed data.
-
-In @fig-ppcheckandposteriorvsprior, we compare the posterior distributions with the prior distributions.
-The comparison shows how the data has influenced the model’s parameters, providing insight into the extent to which our priors were updated by the observed data.
-
-```
-
-#Assumption Checks
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-# Residual vs Fitted values
-plot(attendance_model, which = 1)
-
-# Q-Q plot for normality of residuals
-plot(attendance_model, which = 2)
-```
-
-#Model Validation
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-set.seed(2024)
-train_index <- createDataPartition(cleaned_data$attendance, p = 0.8, list = FALSE)
-train_data <- cleaned_data[train_index, ]
-test_data <- cleaned_data[-train_index, ]
-
-attendance_model <- lm(attendance ~ home_team_id + away_team_id + home_team_score_margin + away_team_score_margin, data = train_data)
-```
 
@@
 
+# Additional Data Details
+
+The English Women's Football (EWF) Database provided detailed match, attendance, and standings data.
+This database includes metrics such as match outcomes, goals for and against, attendance, and standings, which were all used in the analysis.
+Data cleaning and filtering steps were performed to ensure consistency and reliability in the findings presented.
+
+## Model Details and Validation {#sec-model-details}
+
+### Posterior Predictive Check
+A posterior predictive check was conducted to evaluate the extent to which the model captures the patterns in the observed data.
+In @fig-ppcheckandposteriorvsprior, the observed data is compared against replicated datasets generated from the posterior distribution of the model parameters.
+This diagnostic provides insights into areas where the model may underfit or overfit the data.
+
+Additionally, a comparison of posterior distributions with prior distributions is included.
+This comparison highlights how the data has influenced the model’s parameters, illustrating the extent to which prior beliefs have been updated by observed evidence.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-ppcheckandposteriorvsprior
+
+# Posterior predictive check
+# pp_check(attendance_model_data, fun = "hist")
+```
+
+### Assumption Checks
+Diagnostic plots were generated to check for the assumptions of linear regression, @fig-Assumption including:
+
+**Residuals vs. Fitted Values:**
+- This plot checks the linearity assumption by examining whether residuals have constant variance and are centered around zero.
+
+**Q-Q Plot for Normality:**
+- This plot assesses whether residuals are normally distributed, a critical assumption for inference in linear regression models.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-Assumption
+# Residual vs Fitted values
+plot(attendance_model, which = 1)
+
+# Q-Q plot for normality of residuals
+plot(attendance_model, which = 2)
+```
+
+### Model Validation
+
+The dataset was split into training (80%) and testing (20%) sets to validate the model's performance on unseen data.
+This approach ensures that the model generalizes well and is not overfitted to the training data, see @tbl-model-validation.
+
+1. **Train-Test Split:**
+
+- Training data: Used to fit the model.
+- Testing data: Used to evaluate the model's predictive accuracy.
+
+2. **Performance Metrics:**
+
+- Root Mean Squared Error (RMSE): Measures the average prediction error in the same units as the response variable (attendance).
+- Mean Absolute Error (MAE): Captures the average absolute prediction error.
+- R-Squared ($R^2$) Indicates the proportion of variance in attendance explained by the predictors.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: tbl-model-validation
+# Train-test split
+set.seed(2024)
+train_index <- createDataPartition(cleaned_data$attendance, p = 0.8, list = FALSE)
+train_data <- cleaned_data[train_index, ]
+test_data <- cleaned_data[-train_index, ]
+
+# Model fitting
+attendance_model <- lm(attendance ~ home_team_id + away_team_id + home_team_score_margin + away_team_score_margin, data = train_data)
+
+# Predictions on test data
+predictions <- predict(attendance_model, newdata = test_data)
+
+# Performance metrics
+rmse <- sqrt(mean((predictions - test_data$attendance)^2, na.rm = TRUE))
+mae <- mean(abs(predictions - test_data$attendance), na.rm = TRUE)
+r_squared <- summary(attendance_model)$r.squared
+
+metrics <- data.frame(
+  Metric = c("Root Mean Squared Error (RMSE)", 
+             "Mean Absolute Error (MAE)", 
+             "R-Squared (R2)"),
+  Value = c(rmse, mae, r_squared)
+)
+
+# Display the table
+knitr::kable(
+  metrics,
+  caption = "Performance metrics for the linear model predicting attendance."
+)
+```
+### Residual Analysis:
+- Residuals plotted against each predictor to identify non-linear relationships or heteroscedasticity.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+par(mfrow = c(2, 2))
+plot(attendance_model)
+```
+
 \newpage

### 2024-12-15 paper/paper.qmd
Update intro, data section and references

@@
   - Rayan Awad Alim
-thanks: "Code and data are available at: [https://github.com/RayanAlim/EnglishWomensFootballAnalysis/](https://github.com/RayanAlim/EnglishWomensFootballAnalysis)."
+thanks: "Code and data are available at: [github.com/RayanAlim/EnglishWomensFootballAnalysis/](https://github.com/RayanAlim/EnglishWomensFootballAnalysis)."
 date: December 3, 2024
 date-format: long
-abstract: "This paper analyzes attendance and performance trends in the Women's Super League using data from the English Women's Football (EWF) Database.
-Using linear and logistic regression models, we identify that match attendance has steadily increased over time, influenced by major international events.
-Additionally, match attendance correlates with home team success, indicating a measurable 'home advantage.' These insights are crucial for improving team strategies, optimizing league operations, and enhancing fan engagement."
+abstract: "This paper analyzes attendance and performance trends in the Women’s Super League (WSL) using data from the English Women's Football (EWF) Database.
+We examine the evolution of attendance over time and evaluate how key factors—such as goals scored, goals conceded, and league tier—influence both match-level outcomes and season-level team performance.
+Using regression modeling, we find that match attendance has grown substantially, correlating with higher probabilities of home-team success, suggesting a measurable 'home advantage.' These insights provide actionable information for enhancing fan engagement strategies and guiding team development, especially as the intrest in womens sports increases globally."
 format: pdf
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 
-The Women's Super League (WSL) has seen significant growth since its inception, both in terms of attendance and popularity.
-This paper aims to explore trends in WSL attendance and identify key performance factors that determine match outcomes, using the English Women's Football (EWF) Database.
-Understanding these trends is essential for teams, analysts, and league organizers seeking to enhance fan engagement and improve team performance.
+Women’s professional football in England (and globally) has experienced substantial growth in recent years, notably reflected in the Women’s Super League (WSL).
+Increased media coverage, enhanced professionalism, and international success of national teams have all contributed to rising attendance and interest @vanEekeren2022 @Kitching2022.
+Understanding the determinants of match attendance and performance outcomes is valuable for clubs, team analysts, league organizers, and stakeholders seeking to enhance fan engagement, team performances and competitive standards as demonstrated by @GarciaRodriguez2002, attendance determinants can be significantly influenced by local factors.
 
-The estimand of this study is to determine how factors like attendance, team strength, and historical performance affect match outcomes in the Women's Super League.
-By quantifying these relationships, we aim to provide insights into what drives successful match results and how attendance figures have evolved over time.
+This paper explores two central themes: the evolution of match attendance over time and the relationship between key performance indicators (e.g., goals scored, goals conceded) and both match-level outcomes (win, draw, loss) and season-level performance (final standings, points accrued).
+Using the English Women’s Football (EWF) Database(@EWF)- a historical record of top-tier and second-tier women’s football in England- we analyze how attendance trends correlate with on-field results and whether home teams benefit from larger crowds.
 
-The analysis finds that attendance has generally increased over time, with certain key events serving as catalysts for spikes in crowd size.
-Additionally, higher attendance is correlated with an increased likelihood of home team success, suggesting a potential impact of crowd support on team performance.
+The estimand is to understand how factors like attendance, goals, and tier level influence match results and team points.
+Our findings reveal that attendance has steadily increased, especially around peak events and growth phases of the league.
+Higher attendance correlates with a greater likelihood of home-team success, suggesting a measurable “home advantage.” Additionally, goals scored and conceded are strong predictors of a team’s season-long performance.
+The analysis also finds that attendance has generally increased over time, with certain key events serving as catalysts for spikes in crowd size.
 
-This research matters because it helps identify the factors that contribute to successful outcomes in women’s football, providing valuable information for teams to improve strategies, for organizers to boost engagement, and for analysts interested in understanding sports dynamics.
-These findings are particularly significant as women's sports experience unprecedented growth globally, with leagues and federations seeking sustainable ways to attract and retain audiences.
-This study contributes to the growing body of literature that underscores the economic and cultural value of women's football.
+This research matters because it provides insights into what drives successful outcomes and robust fan engagement in women’s football.
+It informs strategic decision-making for clubs aiming to improve both their on-field performances and the experiences they offer to supporters, for organizers to boost engagement, and for analysts interested in understanding sports dynamics.
+More broadly, these findings are particularly significant as women's sports experience unprecedented growth globally,this study contributes to the growing body of literature that underscores the economic and cultural value of women's football where robust data-driven evidence can shape policy, marketing, and development strategies.
 
@@
 
-## Overview
+## Dataset Context and Variable Overview
 
-We use the statistical programming language R [@citeR] to perform our analysis.
-Our data is derived from the English Women's Football (@EWF) Database, which provides a comprehensive dataset of matches, team appearances, and standings in the Women's Super League and Women's Championship.
-Following the guidance provided by @tellingstories, we consider how best to prepare and use these data for analysis in order to effectively tell a story of attendance and performance trends.
-The analysis was conducted using the statistical programming language R [@citeR] and several libraries, including `tidyverse` [@tidyverse], `janitor` [@janitor], `knitr` [@knitr], `dplyr` [@dplyr], `arrow` [@arrow], and `here` [@here] for data manipulation, as well as `ggplot2` [@ggplot], and `kableExtra` [@kableExtra] for visualization.
+Our data is derived from the English Women's Football (@EWF) Database, which provides a detailed dataset of matches, team appearances, and standings in the Women's Super League(tier-1) and Women's Championship(tier-2) in England.
+The dataset includes records spanning multiple seasons, capturing key variables such as attendance, team results, player appearances, and seasonal standings.
+By analyzing these variables, we can better understand the factors driving attendance, match outcomes, and overall team performance in the league.Following the guidance provided by @tellingstories, we considered how best to prepare and use this data for analysis in order to effectively tell a story of attendance and performance trends.
+The analysis was conducted using the statistical programming language R [@citeR] and several libraries, including `tidyverse` [@tidyverse], `ggplot2` [@ggplot], `knitr` [@knitr], `arrow` [@arrow], and `here` [@here] for data manipulation and visualization, as well as  `janitor` [@janitor] and `kableExtra` [@kableExtra] for data cleanaing and presentation.
 
@@
 
-1. *ewf_matches*: Contains all matches played with details like attendance, score, and outcomes.
+1. *`ewf_matches`*: Contains all matches played with details like date, attendance, team and scores.
 
-2. *ewf_appearances*: Contains information about team appearances in each match.
+2. *`ewf_appearances`*:  Contains team-level information for each match, linking teams to their goals and results.
 
-3. *ewf_standings*: Contains end-of-season standings for each team.
+3. *`ewf_standings`*: Contains end-of-season standings, including points, goals for/against, and final positions.
 
-## Measurement
+While other datasets, such as FIFA or Opta's proprietary data, exist, they are not as accessible or as granular at the domestic match level.
+The EWF dataset is chosen for its open availability, rich detail, and focus on the English domestic league structure.
+This level of granularity (per-match attendance and scores, historical standings) is critical to the research questions here, making the EWF dataset the most suitable for our study.
 
-The process of measurement involves translating real-world events into structured data entries.
-In the context of our analysis, this means taking observable phenomena such as the attendance of a football match, the outcome of a game, and individual team performances and converting them into numerical or categorical data points.
+## Measurement and Variable Constructions
 
-For example, attendance is recorded as the number of spectators present at each match.
+Each match played in the Women’s Super League translates into one or more structured entries in the dataset.
+For instance, a match between Arsenal and Chelsea that took place on March 1, 2022, with a recorded attendance of 3,500 spectators and a final score of Arsenal 2 – Chelsea 1, becomes a record in the EWF matches file.
+The attendance figure is typically taken from official league sources, ensuring reliable measurement.
+The goals for each team are recorded from match officials’ reports.
+Similarly, the final league standings at the end of the season, which reflect teams’ cumulative performance and points, are aggregated from individual match results.
+This chain of measurement ensures that every number in the dataset corresponds to a specific, verifiable real-world event.
+
+*Attendance* is recorded as the number of spectators present at each match.
 This variable captures the level of audience engagement and is an indicator of the popularity of the match.
@@
 
-Similarly, match outcomes are recorded as categorical variables, with values such as “Home Win,” “Away Win,” and “Draw.” These categories are derived directly from match results and are used to evaluate performance trends.
-This structured approach allows us to quantify and model the likelihood of different outcomes based on a variety of predictors.
+*Match outcomes* are recorded as categorical variables, with values such as “Home Win,” “Away Win,” and “Draw.” These categories are derived directly from match results and are used to evaluate performance trends.
+These constructed variables allow easier interpretation of offensive and defensive strength.
+
+*Score Margins* (home_team_score_margin, away_team_score_margin) are constructed by subtracting the opponent’s goals from a team’s goals in a specific match.
+This transformation provides a numeric sense of dominance or competitiveness in a game.
 
@@
 
-## Dataset Context and Variable Overview
-
-The EWF Database offers extensive historical data on women's football in England.
-The dataset includes records spanning multiple seasons of the Women's Super League, capturing key variables such as attendance, team results, player appearances, and seasonal standings.
-By analyzing these variables, we can better understand the factors driving attendance, match outcomes, and overall team performance in the league.
-
-While there are other datasets that track women's football, such as those from FIFA or Opta, the EWF dataset was selected due to its specific focus on English domestic competitions and its detailed match-level information.
-Other datasets either lacked granularity or were not publicly accessible, making the EWF dataset the most suitable for our study.
-
 ### Variables of Interest
 
-1. Match-Level Variables (ewf_matches)
+1. *Season and Tier Variables:*
 
-- Season ID (season_id): A unique identifier for each season, allowing us to differentiate between different periods in the dataset.
+- Season ID (season_id): Identifies the season in a format like “2021–2022”. It connects matches to a specific timeframe and competition format. This variable helps track temporal trends.
 
-- Season (season): Represents the year(s) of each season. This variable helps track temporal trends in attendance and performance.
+- Tier (tier): Indicates whether the match was played in the top tier (1) or second tier (2).
+This classification captures differences in competition level, resources, and potentially attendance patterns.
 
-- Tier (tier): Indicates the level of the competition, with the Women's Super League generally representing the highest tier.
-Differences in performance and attendance across tiers are captured through this variable.
+2. *Match Identifiers and Structure:*
 
-- Division (division): Provides additional categorization of the league (e.g., "FA Women's Super League (WSL)").
-
-- Match ID (match_id): A unique identifier for each match.
-
-- Match Name (match_name): The competing teams in each match.
+- Match ID (match_id) and Match Name (match_name): : Unique identifiers specifying the teams involved. These ensure each match is distinguishable and can be cross-referenced with other variables.
 
 - Date (date): The date the match was played. This variable is important for tracking seasonal and temporal trends.
-
-- Attendance (attendance): The number of spectators present at each match. Missing values were filtered out to ensure data consistency.
 
@@
 
-- The outcome variables of interest in this study include match attendance and match outcomes. To understand these outcomes, we provide graphical and tabular representations.
+3. *Attendance Metrics*
 
-### Attendance
-
-The attendance variable captures the number of spectators present at each match, serving as a proxy for fan engagement and the overall popularity of the Women's Super League.
+- Attendance (attendance): The number of spectators present at each match.
+No substantial transformations were needed, although matches without reported attendance were excluded from attendance-specific analyses.The attendance variable captures the number of spectators present at each match, serving as a proxy for fan engagement and the overall popularity of the Women's Super League.
 Analyzing attendance trends not only provides insights into the growth trajectory of women's football but also highlights key moments that have driven fan interest.
-Figure @fig-attendance-over-time illustrates the evolution of attendance over time, uncovering critical shifts and spikes in audience behavior, we can see from @fig-attendance-distribution and @tbl-attendance-summary that attendance is most around the 1000-2000 atendees range.
+Figure @fig-attendance-over-time illustrates the evolution of attendance over time, which shows key shifts and spikes in audience behavior, we can see from @fig-attendance-distribution and @tbl-attendance-summary that attendance is most around the 1000-2000 atendees range.
 The notable spikes during certain seasons, often corresponding with major international tournaments or high-profile matches.
@@
 @tbl-attendance-summary shows the summary of attendance.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-#| label: tbl-attendance-summary
-#| tbl-cap: "Summary statistics of match attendance in Women's Super League."
-
-attendance_summary <- cleaned_data %>%
-  summarise(
-    Total_Matches = n(),
-    Avg_Attendance = mean(attendance, na.rm = TRUE),
-    Median_Attendance = median(attendance, na.rm = TRUE),
-    Max_Attendance = max(attendance, na.rm = TRUE),
-    Min_Attendance = min(attendance, na.rm = TRUE)
-  )
-
-knitr::kable(
-  attendance_summary,
-  caption = "Summary statistics of match attendance in the Women's Super League"
-)
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-#| label: fig-attendance-over-time
-#| fig-cap:  "The trend in match attendance over time in the Women's Super League has increased overall with notable spikes often corresponding with major international tournaments or high-profile matches"
-
-cleaned_data %>%
-  filter(!is.na(attendance)) %>%
-  ggplot(aes(x = as.Date(date), y = attendance)) +
-  geom_line(color = "blue") +
-  labs(
-    title = "Attendance Trends Over Time",
-    x = "Date",
-    y = "Attendance"
-  ) +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5))
-```
 
@@
 #| message: false
-ewf_standings %>%
-  ggplot(aes(x = goals_for, y = points)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
+#| label: tbl-attendance-summary
+#| tbl-cap: "Summary statistics of match attendance in Women's Super League."
+
+attendance_summary <- cleaned_data %>%
+  summarise(
+    "Total Matches Played" = n(),
+    "Average Attendance" = mean(attendance, na.rm = TRUE),
+    "Median Attendance" = median(attendance, na.rm = TRUE),
+    "Highest Attendance" = max(attendance, na.rm = TRUE),
+    "Lowest Attendance" = min(attendance, na.rm = TRUE)
+  )
+
+knitr::kable(
+  attendance_summary,
+  caption = NULL
+)
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-attendance-over-time
+#| fig-cap:  "The trend in match attendance over time in the Women's Super League has increased overall with notable spikes often corresponding with major international tournaments or high-profile matches"
+
+cleaned_data %>%
+  filter(!is.na(attendance)) %>%
+  ggplot(aes(x = as.Date(date), y = attendance)) +
+  geom_line(color = "blue") +
   labs(
-    title = "Relationship Between Goals Scored and Points Earned",
-    x = "Goals Scored",
-    y = "Points"
+    title = "Attendance Trends Over Time",
+    x = "Date",
+    y = "Attendance"
   ) +
@@
 
-### Match Outcomes
+4. *Match Outcomes*
 
-The match outcomes variable is a categorical outcome representing whether the home team won, the away team won, or if the match ended in a draw.
+The match *outcomes* variable is a categorical outcome representing whether the home team won, the away team won, or if the match ended in a draw.
 This outcome helps to assess the impact of various predictors, such as attendance, on the likelihood of different results.
@@
 @tbl-match-outcomes-summary summarizes the average attendance and total matches for each match outcome in the Women's Super League, showing that there games where home team won, had more audience.
+@tbl-goal-statistics provides summary statistics of attendance, showing the minimum, median, mean, and maximum values across all recorded matches.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: tbl-goal-statistics
+#| tbl-cap: "Summary statistics for goals for and points across all seasons"
+
+gp_summary <- ewf_standings %>%
+  summarise(
+    "Mean Goals For" = mean(goals_for, na.rm = TRUE),
+    "Median Goals For" = median(goals_for, na.rm = TRUE),
+    "Mean Points" = mean(points, na.rm = TRUE),
+    "Median Points" = median(points, na.rm = TRUE)
+  )
+
+knitr::kable(gp_summary,
+             caption = NULL)
+```
 
@@
 ```{r}
-#| label: tbl-match-outcomes-summary
-#| tbl-cap: "Summary statistics for match outcomes by attendance and other predictors."
 #| echo: false
@@
 #| message: false
+#| label: tbl-match-outcomes-summary
+#| tbl-cap: "Summary statistics for match outcomes by attendance and other predictors."
 
@@
   match_outcome_summary,
-  caption = "Summary statistics for match outcomes by attendance and other predictors.",
+  caption = NULL,
   col.names = c("Match Outcome", "Average Attendance", "Total Matches")
@@
 ```
+
 # Model {#sec-model}

### 2024-12-15 paper/paper.qmd
Update Paper

@@
 
-The goal of our modeling strategy is to understand how match characteristics affect both attendance and match outcomes. We employ two types of regression models:
+The primary goal of our modeling is to quantify the relationships between match-level factors—such as attendance, competitiveness (score margins), and tier—and key outcomes (match attendance levels, match results, and team points).
+Based on the data described previously, we emploed three models:
 
-1. Linear Regression to Predict Match Attendance.
+1. *Linear Regression Model for Match Attendance:*  
+   To understand which factors influence the number of spectators.
 
-2. Logistic Regression to Predict Match Outcomes (Home Win Probability).
+2. *Logistic Regression Model for Match Outcomes:*  
+   To estimate the probability of a home win, incorporating attendance and performance indicators.
+
+3. *Linear Regression Model for Team Points:*
+   To link season-long performance metrics (goals scored and conceded) to overall success (points earned).
 
 These models allow us to quantify the impact of factors such as goals scored, goals conceded, and attendance on match success and fan engagement.
-We also run the models in R [@citeR] using the `rstanarm` package of [@rstanarm].
-We use the default priors from `rstanarm`.
+All models were fitted using R [@citeR] and the `rstanarm` package for Bayesian regression with default priors [@rstanarm].
 
@@
 Logistic regression for match outcomes appropriately models a binary response variable, allowing us to estimate the probability of a specific result (home win).
+While the default priors are used here for simplicity, these could be refined if domain expertise suggested particular parameter distributions.
 
@@
 Including these features reflects the aspects discussed in the data section, where attendance serves as a proxy for crowd support, and score margins indicate competitive balance.
-While logistic regression was ultimately chosen for predicting match outcomes, alternative methods such as Poisson regression explored for their suitability in modeling count-based phenomena like goals scored.
-However, logistic regression was better for its interpretability and alignment with binary outcomes.
+While logistic regression was ultimately chosen for predicting match outcomes, alternative methods such as Poisson regression explored for their suitability in modeling count-based phenomena like goals scored so it was not chosen due to the main interest being on outcomes and points rather than raw goal counts.
+Also hierarchical models could account for team-level variability, but would add complexity.
+The current approach focuses on fixed effects for clarity and interpretability.
+By considering these alternatives and justifying our chosen models, we show that the final approach is both reasonable and aligned with the research questions and the nature of the data.
 
@@
 
-- Linearity: The linear regression models assume a linear relationship between the predictors and the response. This may not fully capture the complex dynamics of sports events.
+- **Linearity:** The linear models assume a linear relationship between predictors and the outcome. Non-linearities, if present, are not modeled here, but could be explored if diagnostics suggest it.
 
-- Independence: The models assume that observations are independent, which may not hold if there are underlying correlations (e.g., team performance across matches).
+- **Independence of Observations:** Matches involving the same teams might be correlated.
+We do not model random effects here for simplicity, but this could be considered if team-level clustering emerges as important.
 
-- Error Normality: The linear models assume that error terms are normally distributed.
+- **Distributional Assumptions:** For linear models, we assume normally distributed residuals. The logistic model assumes a Bernoulli outcome with a logit link. These are standard assumptions.
 
@@
 
-- Training and Testing Split: The data was split into training and testing sets to evaluate model performance.
+Found in the Appendix :
 
-- Model Diagnostics: Residual analysis and diagnostic plots were used to check for linear regression assumptions. For logistic regression, goodness-of-fit was assessed using measures like deviance.
+- **Data Splits:** The dataset is divided into training and test sets.
+Models are fitted on training sets, and performance metrics (RMSE for linear models, accuracy/log-loss for logistic models) are evaluated on test sets.
 
-- Validation Metrics: Root Mean Square Error (RMSE) was used for linear models, and accuracy was used for logistic regression to evaluate predictive performance.
+- **Diagnostics:** Residual plots, posterior predictive checks, and chain convergence diagnostics (for Bayesian fitting) ensure model appropriateness.
+
+- **Sensitivity Analyses:** Alternative specifications (e.g., dropping outliers or using different prior strengths) may be explored to assess robustness.
 
@@
 
-## Understanding the Factors that Drive Match Attendance {#sec-first-point}
+## Big events Understanding the Factors that Drive Match Attendance {#sec-first-point}
 
-This paper explores factors influencing match attendance and performance in the Women's Super League.
 One of the key findings is that attendance has generally increased over time, with major events and international tournaments serving as pivotal moments that boost interest in women's football.
@@
 # Posterior predictive check
-# pp_check(attendance_model_data, fun = "hist")
+#pp_check(attendance_model, fun = "hist")
 ```
@@
 ### Assumption Checks
-Diagnostic plots were generated to check for the assumptions of linear regression, @fig-Assumption including:
+Diagnostic plots were generated to check for the assumptions of linear regression including:
 
 **Residuals vs. Fitted Values:**
-- This plot checks the linearity assumption by examining whether residuals have constant variance and are centered around zero.
+- This plot checks the linearity assumption by examining whether residuals have constant variance and are centered around zero @fig-res.
 
 **Q-Q Plot for Normality:**
-- This plot assesses whether residuals are normally distributed, a critical assumption for inference in linear regression models.
+- This plot assesses whether residuals are normally distributed, a critical assumption for inference in linear regression models @fig-qq.
 
@@
 #| message: false
-#| label: fig-Assumption
+#| label: fig-res
+#| fig-cap: "Diagnostic plot for linear regression assumptions: The Residuals vs.
+Fitted Values plot checks for linearity and homoscedasticity by assessing whether residuals are randomly dispersed around zero with constant variance."
+
 # Residual vs Fitted values
 plot(attendance_model, which = 1)
+```
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-qq
+#| fig-cap: "Diagnostic plots for linear regression assumptions: The Q-Q plot evaluates the normality of residuals, ensuring they follow a standard normal distribution critical for valid inference in linear regression."
 
@@
 #| label: tbl-model-validation
+#| tbl-cap: "Performance metrics for the linear model predicting attendance."
+
 # Train-test split
@@
 knitr::kable(
-  metrics,
-  caption = "Performance metrics for the linear model predicting attendance."
+  metrics
 )
 ```
+
+\newpage
+
 ### Residual Analysis:
-- Residuals plotted against each predictor to identify non-linear relationships or heteroscedasticity.
+- Residuals plotted against each predictor to identify non-linear relationships or heteroscedasticity @fig-resid.
 
@@
 #| message: false
+#| label: fig-resid
+#| fig-cap: "Residuals plotted against each predictor to identify non-linear relationships or heteroscedasticity."
+
 par(mfrow = c(2, 2))

### 2024-12-15 paper/paper.qmd
update paper

@@
 We examine the evolution of attendance over time and evaluate how key factors—such as goals scored, goals conceded, and league tier—influence both match-level outcomes and season-level team performance.
-Using regression modeling, we find that match attendance has grown substantially, correlating with higher probabilities of home-team success, suggesting a measurable 'home advantage.' These insights provide actionable information for enhancing fan engagement strategies and guiding team development, especially as the intrest in womens sports increases globally."
+Using regression modeling, we find that match attendance has grown substantially, correlating with higher probabilities of home-team success, suggesting a measurable 'home advantage.' These findings provide actionable information for enhancing fan engagement strategies and guiding team development, especially as the intrest in womens sports increases globally."
 format: pdf
@@
 Increased media coverage, enhanced professionalism, and international success of national teams have all contributed to rising attendance and interest @vanEekeren2022 @Kitching2022.
-Understanding the determinants of match attendance and performance outcomes is valuable for clubs, team analysts, league organizers, and stakeholders seeking to enhance fan engagement, team performances and competitive standards as demonstrated by @GarciaRodriguez2002, attendance determinants can be significantly influenced by local factors.
+Understanding the determinants of match attendance and performance outcomes asissts clubs, team analysts, league organizers, and stakeholders seeking to enhance fan engagement, team performances and competitive standards as demonstrated by @GarciaRodriguez2002, attendance determinants can be significantly influenced by local factors.
 
@@
 
-This research matters because it provides insights into what drives successful outcomes and robust fan engagement in women’s football.
+This research matters because it provides an outlook into what leads to successful outcomes and robust fan engagement in women’s football.
 It informs strategic decision-making for clubs aiming to improve both their on-field performances and the experiences they offer to supporters, for organizers to boost engagement, and for analysts interested in understanding sports dynamics.
-More broadly, these findings are particularly significant as women's sports experience unprecedented growth globally,this study contributes to the growing body of literature that underscores the economic and cultural value of women's football where robust data-driven evidence can shape policy, marketing, and development strategies.
+More broadly, these findings are particularly significant as women's sports experience unprecedented growth globally,this study contributes to the growing body of literature that underscores the economic and cultural value of women's football where robust data based evidence can shape policy, marketing, and development strategies.
 
@@
 The EWF dataset is chosen for its open availability, rich detail, and focus on the English domestic league structure.
-This level of granularity (per-match attendance and scores, historical standings) is critical to the research questions here, making the EWF dataset the most suitable for our study.
+This level of granularity (per-match attendance and scores, historical standings) is needed for the research questions here, making the EWF dataset the most suitable for our study.
 
@@
 - Home Team Score Margin (home_team_score_margin) and Away Team Score Margin (away_team_score_margin): Calculated as the difference between goals scored by the respective team and the opponent.
-These variables provide insight into match competitiveness.
+These variables provide an understanding into match competitiveness.
 
@@
 No substantial transformations were needed, although matches without reported attendance were excluded from attendance-specific analyses.The attendance variable captures the number of spectators present at each match, serving as a proxy for fan engagement and the overall popularity of the Women's Super League.
-Analyzing attendance trends not only provides insights into the growth trajectory of women's football but also highlights key moments that have driven fan interest.
+Analyzing attendance trends not only provides us information about the growth trajectory of women's football but also highlights key moments that have driven fan interest.
 Figure @fig-attendance-over-time illustrates the evolution of attendance over time, which shows key shifts and spikes in audience behavior, we can see from @fig-attendance-distribution and @tbl-attendance-summary that attendance is most around the 1000-2000 atendees range.
@@
 For instance, attendance saw a significant increase following the FIFA Women's World Cup, indicating a spillover effect where global events bolster domestic league interest.
-This suggests that leveraging international exposure through strategic marketing and scheduling can have a profound impact on league engagement.
+This suggests that leveraging international exposure through strategic marketing and scheduling can have a huge impact on league engagement.
 @tbl-attendance-summary shows the summary of attendance.
@@
 
-The analysis reveals that factors such as the teams playing and the score margin (both home and away) are significant predictors of match attendance.
+The analysis highlights that factors such as the teams playing and the score margin (both home and away) are significant predictors of match attendance.
 Popular teams and closely contested matches tend to attract more spectators.
@@
 The logistic regression results show that higher attendance significantly increases the probability of a home team victory, suggesting the influence of crowd support.
-Furthermore, the score margins for both teams are critical factors in determining match outcomes.
+Furthermore, the score margins for both teams are important factors in determining match outcomes.
 
@@
 One of the key findings is that attendance has generally increased over time, with major events and international tournaments serving as pivotal moments that boost interest in women's football.
-Understanding what drives this interest allows league organizers to align promotional efforts with these catalysts.
+Understanding what is the basis of this interest allows league organizers to align promotional efforts with these catalysts.
 
@@
 
-## Both offensive and defensive abilities are crucial for a team's performance
+## Both offensive and defensive abilities are key for a team's performance
 
@@
 Goals scored positively affect match outcomes, as expected, while goals conceded correlate negatively with team success.
-These results confirm the intuitive idea that both offensive and defensive abilities are crucial for a team's performance.
+These results confirm the intuitive idea that both offensive and defensive abilities are key for a team's performance.
 It is not enough to simply score goals-preventing the opponent from scoring also plays an important role.
@@
 
-Moreover, the analysis of attendance reveals that larger crowds tend to coincide with better performances for the home team.
+Moreover, the analysis of attendance shows us that larger crowds tend to coincide with better performances for the home team.
 This finding highlights the potential advantage that crowd support can provide, supporting the concept of "home-field advantage" in sports.
@@
 
-For organizers, these insights are valuable for shaping promotional strategies that increase attendance, leveraging key calendar events, and enhancing audience experience, thereby boosting overall league engagement and team performance.
+For organizers, these findings help in shaping promotional strategies that increase attendance, leveraging key calendar events, and enhancing audience experience, thereby boosting overall league engagement and team performance.
 
@@
 Another limitation is the absence of player-specific performance metrics.
-Incorporating individual-level data, such as player fatigue or injuries, would add depth to our understanding of what drives match outcomes.
+Incorporating individual-level data, such as player fatigue or injuries, would add depth to our understanding of what affects match outcomes.
 Future models could incorporate additional predictors to improve predictive power and capture the complex dynamics of match performance.
@@
   While this study does not directly use survey data, similar analyses in the literature often rely on surveys to capture fan demographics, motivations, and preferences.
-Incorporating such surveys into future research could provide richer contextual insights.
+Incorporating such surveys into future research could provide richer contextual understanding
 
@@
 In @fig-ppcheckandposteriorvsprior, the observed data is compared against replicated datasets generated from the posterior distribution of the model parameters.
-This diagnostic provides insights into areas where the model may underfit or overfit the data.
+This diagnostic highlights potential areas where the model might underperform or overfit the data.
 
@@
 #| label: fig-qq
-#| fig-cap: "Diagnostic plots for linear regression assumptions: The Q-Q plot evaluates the normality of residuals, ensuring they follow a standard normal distribution critical for valid inference in linear regression."
+#| fig-cap: "Diagnostic plots for linear regression assumptions: The Q-Q plot evaluates the normality of residuals, ensuring they follow a standard normal distribution for valid inference in linear regression."
 
